### 2023-07-19 article/article.tex
adding description of default TeX rules

% xelatex --shell-escape article.tex 
% pandoc -s article.tex --pdf-engine=xelatex --pdf-engine-opt=-shell-escape -o article.html
\documentclass{article}
\usepackage[english]{babel}
%\usepackage{testhyphens}
\usepackage{multicol}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{longtable}
\usepackage[style=alphabetic, backend=biber]{biblatex}
\addbibresource{references.bib}
\usepackage[inline]{enumitem}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{fancyvrb}
\usepackage{siunitx}

%%%%%%% https://tex.stackexchange.com/questions/16790/write18-capturing-shell-script-output-as-command-variable
\usepackage{xparse}
\ExplSyntaxOn
\NewDocumentCommand{\captureshell}{som}
 {
  \sdaau_captureshell:Ne \l__sdaau_captureshell_out_tl { #3 }
  \IfBooleanT { #1 }
   {% we may need to stringify the result
    \tl_set:Nx \l__sdaau_captureshell_out_tl
     { \tl_to_str:N \l__sdaau_captureshell_out_tl }
   }
  \IfNoValueTF { #2 }
   {
    \tl_use:N \l__sdaau_captureshell_out_tl
   }
   {
    \tl_set_eq:NN #2 \l__sdaau_captureshell_out_tl
   }
 }

\tl_new:N \l__sdaau_captureshell_out_tl

\cs_new_protected:Nn \sdaau_captureshell:Nn
 {
  \sys_get_shell:nnN { #2 } { } #1
  \tl_trim_spaces:N #1 % remove leading and trailing spaces
 }
\cs_generate_variant:Nn \sdaau_captureshell:Nn { Ne }
\ExplSyntaxOff
%%%%%%%

\title{Improving the portuguese hyphenation rules}
\author{Leonardo Araújo \and Aline de Lima Benevides}
\begin{document}
\VerbatimFootnotes
\maketitle

\captureshell*[\NumberOfDefaultRules]{wc -l ../data/default.TeX.pt-br.patterns | cut -d' ' -f1}

\captureshell*[\NumberOfNewRules]{wc -l ../data/patch.TeX.pt-br.patterns | cut -d' ' -f1}

% 6 dictionaries --- 5 commas
\captureshell*[\DictionarySize]{../scripts/dicstats.sh}

% https://www.texdev.net/2009/10/06/what-does-write18-mean/
\captureshell*[\NumberOfSixHyphens]{../scripts/dicstats.sh 6}
\captureshell*[\NumberOfFiveHyphens]{../scripts/dicstats.sh 5}
\captureshell*[\NumberOfFourHyphen]{../scripts/dicstats.sh 4}
\captureshell*[\NumberOfThreeHyphens]{../scripts/dicstats.sh 3}
\captureshell*[\NumberOfTwoHyphens]{../scripts/dicstats.sh 2}
\captureshell*[\NumberOfOneHyphens]{../scripts/dicstats.sh 1}
\captureshell*[\NumberOfNoHyphens]{../scripts/dicstats.sh 0}
% run to update values
%\immediate\write18{../scripts/hyphenationagreements.sh > ../data/hyphagreements}
\captureshell*[\NumberOfSixAgrees]{../scripts/gethyphenationagreements.sh 6}
\captureshell*[\NumberOfFiveAgrees]{../scripts/gethyphenationagreements.sh 5}
\captureshell*[\NumberOfFourAgrees]{../scripts/gethyphenationagreements.sh 4}
\captureshell*[\NumberOfThreeAgrees]{../scripts/gethyphenationagreements.sh 3}
\captureshell*[\NumberOfTwoAgrees]{../scripts/gethyphenationagreements.sh 2}
\captureshell*[\NumberOfOneAgrees]{../scripts/gethyphenationagreements.sh 1}

\newcounter{numberRulesGroups}

\begin{abstract}
    Portuguese hyphenation rules are available for more than 35
    years and have done a good job. Nonetheless they still make mistakes and
    leave some hyphenation points unmarked. Although most undetected
    hyphenation points are located near word boundaries, what will be
    irrelevant for \TeX{} typographic purposes, they are still useful to
    hyphenate proper nouns, new words or pseudoword, and for usage in other
    applications, such as text-to-speech conversion. A list of
    \DictionarySize{} hyphenated words acquired from online dictionaries was
    used along \emph{patgen} to create improved rules, leading to better hyphenation
    of Portuguese words.
\end{abstract}

\section{Introduction}\label{sec-intro}
Hyphenation in text wrapping was not used for a long time. Words should fit
entirely in a line, or they would be broken in arbitrary places. First no
marker at all was used to sign the wrapping of a word, what could create
confusion and undesirable meaning. Therefore, ortographers advocated for the
use of a sign to indicate this break. Portuguese faced the same gradual
introduction of a hyphenation sign to mark words wrapping across lines. Even
though the usage of a hyphenation sign ($=$) was advocated by orthographers
\cite{gandavo1574}, few documents used such sign until the end of the 18th
century \cite{araujo2015}.

In some cases, hyphenation hinders smooth reading and should be avoided in
child literature, for example. In opposition, large or small spaces between
words also impose difficulty in the reading process, making hyphenation
fundamental when texts use short line lengths. As a line gets shorter, the
number of breaking candidates between words decreases, leading to awkward
spaces between words and among letters. For that matter, automatic hyphenation
plays an important role in good typesetting.

%Automatic hyphenation plays an important role in good typesetting and
%it is fundamental when a text uses short line lengths. As a line gets
%shorter, the number of breaking candidates between words decreases,
%creating awkward spaces between words and among letters.

\TeX{} is a typesetting system which carefully deals with these issues,
automatically arranging text on a page to create a good reading experience.
Automatic hyphenation is an important part on this process,
promoting an even-tempered distribution of elements on the page.
%Hyphenation is therefore an important part on this process.
%A good readability depends on an even-tempered distribution of elements
%on the page.
Line height and length, paragraph length, font size and typeface,
letter and word-spacing, are some factors which influence the text legibility
and readability. Space between words should not be too long creating lakes and
rivers in the text, nor too tight, impairing the legibility and readability.

Another important matter to consider is ambiguities that might be created when
a word is partitioned during the hyphenation process. In English we should
avoid hyphenations such as \emph{re-cover}, \emph{re-form}, \emph{re-sign}, \emph{the-rapist},
\emph{depart-mental}, and \emph{mans-laughter} (in Portuguese, some examples are:
\emph{de-putada}, \emph{fede-ração}, \emph{acu-mula}, \emph{após-tolo},
\emph{cú-bico}). Hyphenations that might lead the reader to pronounce a word
incorrectly should also be prevented. That is the case of
\emph{considera-tion}, in Enslish (and \emph{pe-rigo} in Portuguese).

In some situations, hyphenation is also a matter of style. Some partitioning
choices sound better than others. These conflicting alternatives typically
arises when a words has many possible hyphenation points. Consider
\emph{ar-chae-ol-o-gist} (\emph{or ar-che-ol-o-gist}), which is preferably partitioned as
\emph{archae-ologist} (or \emph{arche-ologist}) in opposition to \emph{archaeol-ogist} (or
\emph{archeol-ogist}) or \emph{archaeolo-gist} (or \emph{archeolo-gist}). It is preferable to keep
whole morphemes together. In the previous example: \emph{archae} (or \emph{arche}, meaning
``ancient'', ``primitive'')
%\footnote{archae- comes from the Ancient Greek ἀρχαῖος (arkhaîos, `ancient',
%`primitive'), from ἀρχή (arkhḗ, `beginning').}
%https://en.wiktionary.org/wiki/archae-
and \emph{-ologist} (``one who studies the topic''). In Portuguese, it is also more
elegant to avoid splits between double consonants or vowels, even if an
hyphenation point do exists between those letters. For example,
\emph{pressu-rizar} is preferable over \emph{pres-surizar} and
\emph{empreen-dedor} is preferable over \emph{empre-endedor}. Even so,
exceptions exists, it is preferable to partition \emph{micro-organismo}
rather than \emph{microor-ganismo} (keeping morphemes together is favored
over splitting a double vowel). Numerous factors come into play when choosing
the optimal hyphenation point for a word.

The general rule for Portuguese hyphenation is to split a word into its
syllables. A syllable is made of a mandatory nucleus, filled by a vowel, and
optional peripheral consonants (before or after the nucleus). In some
situations, the syllabic division does not respect the ethnologic constituents.
The usage of the prefixes \emph{bis-} and \emph{in-} are examples of this
circumstance. The correct syllabifications are \emph{bi-sa-vô},
\emph{i-na-ti-vo}, and \emph{i-nobs-tan-te}\footnote{The rule of syllable
division could lead to two possible partitions: \emph{i-nobs-tan-te} and
\emph{in-obs-tan-te}, but the first is preferable.}, where the prefixes are
split into two syllables. But, as pointed previously, it is preferable to keep
morphemes together rather than splitting them apart, therefore we should
favor the hyphenation \emph{bisa-vô} over \emph{bi-savô}, and \emph{ina-tivo}
over \emph{i-nativo}. This last example could also lead to a misunderstanding
since the word \emph{nativo} (\emph{native} in English) emerges from this word
break.

Each language has its own hyphenation rules. They might be grouped into two
groups: those which hyphenation is driven by morphology (etymology) and those
driven by pronunciation. An algorithmic approach might use a logic system to
analyse words and apply the hyphenation rules of a given language. An algorithm
must be written for each language, as hyphenation rules vary extensively.
Although a logic system might be fast and compact, inevitably it will have to
deal with exceptions through hard-coded rules.

The automation of this process might use a dictionary based approach, which
will restrict the hyphenation possibilities to those entries in the dictionary,
an algorithmic base approach, which might be applied to whatever sequence found
in a text, or a mix of both approaches. 
%An algorithmic approach might use a logic system to analyse words and apply
%the hyphenation rules of a given language. 
A rule base model might include the recognition of prefixes,
suffixes, morphemes, or some sequences suitable for the inclusion of a break.
Another approach is the usage of pattern matching. Given a corpus of
hyphenation examples in a language, it is necessary to identify those sequence
of letters that define a point where hyphenation is suitable or not. Patterns
might include prefixes, suffixes, exceptions and special hyphenation rules of a
given language. Such approach might be used for different languages, by
providing hyphenation examples in a target language to extract patterns and
rules.

Only the analysis of immediate surrounding might not be enough to determine a
potential break point. For example, consider \emph{dem-o-crat} and \emph{de-moc-ra-cy},
where immediate surrounding of \emph{e} does not suffice to determine the break
location. Even after chosen patterns that makes hyphenation pretty
straightforward, there might be exceptions. Consider the sequence \emph{tion}. It
seems safe to always place a break before this pattern. However, the word
\emph{cation} is hyphenated as \emph{cat-ion}, making is etymology determinant to the way
this word will be split.

A word with different meanings may be hyphenated differently according to its
meaning. `For instance, the Swedish word form \emph{glassko} has three different
meanings, and can be hyphenated as \emph{glas-sko} (glass shoe),
\emph{glass-ko} (ice cream cow) and in the non-standard way,
\emph{glass-sko} (ice cream shoe).' \cite{nemeth2006}. In Portuguese, the
word \emph{sublinha} might be hiphenated in two different forms:
\emph{su-bli-nha} when representing the inflected form of verb
\emph{sublinhar} (underline as a verb) or \emph{sub-li-nha} when refering
to the line under (underline as a noun).

% hyphenation might change with syntactic function
% part of speech (PoS)

Machine learning approach was used to hyphenate Norwegian text
\cite{kristensen2001}. It was found that the \TeX{} approach, in general,
performed better then a neural network. Both approaches had similar
performance on finding the correct hyphenation points and avoiding wrong
hyphenation points on a small word list (the neural network performed slightly
better on finding the correct hyphenation points). Nonetheless, \TeX{} approach overcame
the neural network on a large word list, regarding the avoidance of wrong
hyphenations.

The original \TeX{} hyphenation algorithm was introduced by Knuth 1977
\cite{knuth1977}. It was focused on the English language and used three main
rules:
\begin{enumerate*}[label=(\arabic*)]
 \item suffix removal; 
 \item prefix removal; and 
 \item vowel-consonant-consonant-vowel (VCCV) breaking. 
\end{enumerate*}
Tests showed it could find
40\% of the allowed hyphen locations \cite{liang1983}. 
The hyphenation algorithm proposed by Frank M. Liang and adopted in \TeX{}
utilizes competing patterns. The \TeX{}82 algorithm employs five alternating
levels of hyphenating and inhibiting patterns. The program for pattern
generation, known as PATGEN, was created by Liang \cite{liangbreitenlohner1999}
and has been utilized to generate hyphenating patterns for numerous languages
\cite{sojka1995,sojka1995a,sojka2005thesis,sojka2003,scannell2003}. It involves
sweeping a database of hyphenated words in a language to identify both
hyphenating and inhibiting patterns, ultimately creating a list of competing
patterns for that specific language.

The effective hyphenation of words by \TeX{} will actually depends on the following factors:
\begin{enumerate*}[label=(\arabic*)]
    \item document language, which will determine which set of patterns to apply;
    \item characters used, since some might block hyphenation at their edges;
    \item the value of the internal variables \verb|lefthyphenmin| and \verb|righthyphenmin|\footnote{
	The variables \verb|lefthyphenmin| and \verb|righthyphenmin| are language dependant and
	are defined in \emph{tlpobj} files (\verb|/usr/local/texlive/20XX/tlpkg/tlpobj/hyphen-xxxxxx.tlpobj|). 
	Default values varies in the range from 1 to 3. 
    	English and Portuguese, for example, use \verb|lefthyphenmin=2| and \verb|righthyphenmin=3|.},
	% ls /usr/local/texlive/2023/tlpkg/tlpobj/*.tlpobj | while read -r filename; do grep -Po "lefthyphenmin=[0-9]\s+righthyphenmin=[0-9]" $filename; done | tr '\t' ' ' | tr -s ' ' | sort | uniq -c
	%      17 lefthyphenmin=1 righthyphenmin=1
        %       6 lefthyphenmin=1 righthyphenmin=2
        %       1 lefthyphenmin=1 righthyphenmin=3
        %      49 lefthyphenmin=2 righthyphenmin=2
        %       9 lefthyphenmin=2 righthyphenmin=3
        which defines the minimum sequence length of characters at the left and right borders
        before any hyphenation is allowed.
\end{enumerate*}

Despite the fact that \TeX{} hyphenation algorithm and rules are old, they are,
to these days, the most frequently used approach, even outside the \TeX{}'s
world. The grounds for this is Hunspell, a spell checker and morphological
analyzer that is adopted in many softwares (e.g. LibreOffice, OpenOffice.org,
Mozilla Firefox, Mozilla Thunderbird, Google Chrome, macOS, InDesign, memoQ,
Opera, Affinity Publisher, among others \cite{hunspell}). Hunspell uses \TeX{}
hyphenation rules \cite{hunspellhyphen,levien1998}, making \TeX{} hyphenation
widespread in the computer world. That is a result of \TeX{} approach
simplicity and versatility.  The algorithm works effectively, as it already
supports rules for 66 languages \cite{texhyphenrules}, and offers the
flexibility to create rules for any currently unsupported languages.

Unfortunately, certain hyphenation rules cannot be implemented using the \TeX{}
hyphenation algorithm. In languages like German, hyphenation can lead to letter
changes or insertions. Additionally, compound words lack hyphens, resulting in
extended letter sequences without visible separation and even repetitions of
the same letter, as seen in examples like \emph{Wasserrinne} and
\emph{Schifffahrt}.  Furthermore, the German spelling reform made some changes,
making it necessary to create a different set of rules for German hyphenation.
For example, the word \emph{Schiffahrt} should be hyphenated as
\emph{Schiff-fahrt}, preserving the \emph{f}s from each word that makes this
compound. The hyphenation should insert a \emph{f} that is not part of the
written form. That was not a problem for the old written form of the word:
\emph{Schifffahrt}. Also, the old hyphenating rules of German grammar stated
the hyphenation \emph{Bäk-ker} for the word \emph{Bäcker}, \emph{Zuk-ker} for
the word \emph{Zucker} and \emph{pak-ken} for the word \emph{packen}. Now those
words are hyphenated as \emph{Bä-cker}, \emph{Zu-cker} and \emph{pa-cken},
respectively. Some words have also changed their hyphenation point after the
spelling reform. For example, \emph{Fen-ster} became \emph{Fens-ter} and
\emph{mei-stens} became \emph{meis-tens}. Some problems in compound word
hyphenation in \TeX{} are discussed in \cite{sojka1995a}.

% Some hyphenation rules are not possible with TeX algorithm.
% For example, the german word Schiffahrt should be hyphenated as Schiff-fahrt.
% Although it is no longer the recommend written form, it was the most usual
% until 2003 (see graph in schifffart.png from google ngrams).
% Also the old hyphenating rules of German grammatic stated the following
% hyphenations Bäk-ker (for the word Bäcker, what is now hyphenated as Bä-cker),
% Zuk-ker (for the word Zucker, what is now hyphenated as Zu-cker),
% pak-ken (for packen, and now pa-cken) and trok-ken (for trocken, now tro-cken).

%which contain archeo can be alternatively spelt with archaeo or archæo.

%GÂNDAVO, Pero de Magalhães de. Regras que ensinam a maneira de escrever e
%orthographia da lingoa portuguesa, etc. Lisboa: Antonio Gonsalvez, 1574

%  HIFENIZAÇÃO EM PORTUGUÊS
% Antonio Martins de ARAÚJO and Toru MARUYAMA

\section{Patterns for \TeX{} hyphenation}\label{sec-patt4TeX} 
The patterns used in \TeX{} hyphenation are of the form: \texttt{4z1z2}. Each
pattern is made of a sequence of letters and numbers. Odd numbers indicate a
good hyphenation point, whereas even numbers indicate a bad place to break.
The given example states that the sequence has a good breaking point between
the first and the second \emph{z} and an hyphenation should be inhibited before
the first \emph{z} and after the second \emph{z}. For example, the hyphenation
of the word \emph{piz-za}, \emph{fiz-zle} and \emph{mez-zanine} use this rule,
where we see the hyphen placed between the two \emph{z}'s and no hyphen before,
nor after the \emph{z}'s. The pattern may also use period symbol (\emph{.}) to
indicate word boundaries. The pattern \texttt{.s2h2} applies to beginning of
words, implying that the \emph{s} and the \emph{h} should stick together in
beginning of a word and an hyphenation should also be inhibited after the
\emph{h}. For example, this pattern is used in \emph{Sher-lock}.

Hyphenation rules are organized in levels, from 1 to 9, where odd numbers
represent hyphenating levels and even numbers represent inhibiting levels. Each
level works as an exception level of it predecessor. For example, the rule
\texttt{sh1er} indicate a good hyphenation point between the \emph{h} and the
\emph{e} in the sequence \emph{sher}. A rule at a higher level, as
\texttt{.s2h2}, implies an exception to the lower level rule. When we see
\emph{sher}, in the beginning of a word, the rule \texttt{.s2h2} applies and
hyphenation proposed by the lower level rule \texttt{sh1er} should be hindered.
That is the case in the hyphenation of the word \emph{Sher-lock}. The full
example is provided in \Cref{sherlockhyphenation}, where we might see all
English rules that takes place in the hyphenation of \emph{Sher-lock}.

\begin{lstlisting}[language={}, caption={Example of rules applyied in the
hyphenation of the word \emph{Sherlock}. Example done using a port of \TeX{}'s
hyphenation algorithm to Go provided at
\url{https://github.com/speedata/hyphenation}.}, label=sherlockhyphenation]
   .   s   h   e   r   l   o   c   k   .
     0   0   2   |   |   |   |   |   |    .sh2
     0   2   0   |   |   |   |   |   |    s2h
     0   0   1   0   0   |   |   |   |    sh1er
     |   |   |   0   1   0   |   |   |    r1l
     |   |   |   0   3   0   4   |   |    r3lo4
     |   |   |   |   |   |   0   0   1    ck1
max: 0   2   2   0   3   0   4   0   1
final: s   h   e   r - l   o   c   k -
\end{lstlisting}

A pattern will consist of a string made of characters (from the language
alphabet) possibly with a number in between, expressing the
hyphenation/inhibition level and possibly word boundaries marker (the period)
at the pattern edges. When there is no number between characters in a pattern,
a zero is assumed, which means \emph{undefined} and no hyphenation point will
be suggest at that location.

\section{PATGEN}\label{sec-patgen}

PATGEN utilizes a list of hyphenated words to define rules at various levels
and lengths, based on the patterns found in the data. It starts with short
patterns and incrementally increases their length until reaching the maximum
pattern length allowed by the user.  The objective is to keep the patterns as
concise as possible, as this enhances their generalizability.  As it advances
and incorporates longer patterns, PATGEN establishes exceptions.  In certain
cases, analyzing long patterns may be necessary, as some hyphenation points
could depend on characters far away from the breaking point\footnote{Some
    examples of hyphenation dependency on characters far from the break point:
    \emph{dem-o-crat} and \emph{de-moc-ra-cy}; \emph{as-pi-rin} and
    \emph{aspir-ing}; \emph{de-mon-stra-tive} and \emph{dem-on-stra-tion}.
}.

PATEGEN works on glyph\footnote{\emph{Glyph} is commonly used in linguistics,
typography, and computer graphics to refer to a specific graphical
representation of a character or symbol, which can be the entire symbol or a
distinct visual element within it.} indices rather than character codes. Each
glyph is represented by a single byte. That amount to 256 indices, where 13 of
them are reserved for the digits 0-9 and the characters `.', `-', and `*'. The
remaining 243 are used to represent symbols of the given language.  To run
PATGEN, a translation file is necessary. This file defines the values of
certain language-specific parameters (in the first line) and enumerates the
various forms in which language symbols may appear (all subsequent lines).  In
the first line, positions 1 and 2 are used to set the value of
\verb|lefthyphenmin|, and positions 3 and 4 are used to set the value of
\verb|righthyphenmin|.  These values determine the minimum length of a string
that may be generated by a hyphenation procedure. To set a single-digit value,
leave the first position blank, i.e., place a space in position 1 and 3 for
\verb|lefthyphenmin| and \verb|righthyphenmin|, respectively.  Positions 5, 6
and 7 are used to define alternative values for the special characters `.', `-'
and `*'.  The following lines use a delimiter to enclose each 'letter' of the
desired language alphabet, including its alternative representations.  The
first position of the line defines the delimiter, and each symbol of the
language can occupy as many positions as necessary, as long as the reserved
value for the delimiter is not used in the symbol's definition.  Consider the
following example for defining the letter `e' in Portuguese: 
\begin{verbatim}
XeXEXéXêXÉXÊX\'{e}X\^{e}X\'{E}X\^{E}XX 
\end{verbatim}
It represents the many ways in which the letter `e' might be found: lowercase,
uppercase, with or without acute or circumflex accents. Note that we have used
the direct input (using UTF-8 or other encoding that support the accented `e'
character) and also the compositional counterpart using the escape codes that
define each accent.  In this definition, we have assumed that the many forms in
which we may find the character \emph{e} will be equivalent for
hyphenation (pattern matching) purpose. As another example, see the next line
which defines the character $\pi$ (taken from \cite{haralambous2021}):
\begin{verbatim} 
#p#P#\varpi ## 
\end{verbatim}

The dictionary file is a list of pre-hyphenated words from which PATGEN
extracts patterns to create hyphenation rules. To ensure PATGEN's proper
functionality, the translation file and the dictionary file must utilize the
same encoding, even if it is a multi-byte encoding. The translation file
describes how to handle byte sequences representing a glyph, and PATGEN will
work seamlessly when there are at most 243 symbols in the given language.

%Those rules are organized
%in different levels, from 1 to 9, where odd numbers represent hyphenating
%levels and even numbers represent inhibiting levels. Each level works as an
%inhibiting level of it predecessor. 
%A pattern will consist of a string made of
%characters (from the language alphabet) possibly with a number in between,
%expressing the hyphenation/inhibition level and possibly word boundaries marker
%(the period) at the pattern edges. When there is no number between characters
%in a pattern, a zero is assumed, which means \emph{undefined} and no
%hyphenation point will be suggest at that location.

\section{\TeX{} hyphenation rules for portuguese}\label{sec-tex-hyphen-pt}
\cite{rezende1987} was the first to create patterns for Portuguese hyphenation
in \TeX{}. The pattern set was last updated in 2015, incorporating
contributions from José João Dias Almeida \cite{hyphpt}, resulting in a total
of 307 rules\footnote{The set of rules for Portuguese hyphenation is short when
compared to rules in other languages. For example, English currently has 4938
rules, Russian has 7023 rules and German has 34011 rules.}. 
These rules effectively hyphenate the majority of Portuguese words.

However, in light of certain cases, we propose an analysis of the default rules
to identify areas for improvement. By addressing specific issues and
considering non-typical patterns, we aim to enhance hyphenation accuracy. The
methodology used for this analysis is described in Section
\ref{sec-methodology}.

%
% FILE="../data/default.TeX.pt-br.patterns"; grep "^1..$" "$FILE" | awk '{print substr($0, 0, 2)}' | sort | uniq -c | while read line; do p=$(echo $line | cut -d' ' -f2); c=$(grep "^$p[^0-9]$" "$FILE" | awk '{print substr($0, 3)}' | tr '\n' ' '); printf "$line $c \n"; done | cut -d' ' -f1 | paste -sd+ | bc
%
Out of the default \TeX{} rules, 252 follow the pattern \verb|1CV|, which
represents the recurring CV syllables in Portuguese.  As for the consonants,
there are typically 18 considered (\verb|b|, \verb|c|, \verb|ç|, \verb|d|,
\verb|f|, \verb|g|, \verb|j|, \verb|k|, \verb|l|, \verb|m|, \verb|n|, \verb|p|,
\verb|r|, \verb|s|, \verb|t|, \verb|v|, \verb|x|, and \verb|z|) that can
combine with 14 vowels (\verb|a|, \verb|e|, \verb|i|, \verb|o|, \verb|u|,
\verb|á|, \verb|â|, \verb|ã|, \verb|é|, \verb|í|, \verb|ó|, \verb|ú|, \verb|ê|,
and \verb|õ|), resulting in these CV patterns that indicate favorable
hyphenation points before the consonants.  It is worth noting that there might
be other rules or exceptions in the hyphenation patterns not covered by these
default rules.  To accommodate exceptions, 23 additional rules were created for
cases involving consonants \verb|b|, \verb|c|, \verb|d|, \verb|f|, \verb|g|,
\verb|k|, \verb|p|, \verb|t|, \verb|v|, or \verb|w| followed by \verb|l| or
\verb|r|; or \verb|c|, \verb|l| or \verb|n| followed by \verb|h|.

%
% FILE="../data/default.TeX.pt-br.patterns"; grep -v "^1..$" "$FILE" | grep "2" | awk '{print substr($0, 2, 1)}' | sort | uniq | while read line; doc=$(grep "^1${line}2.$" "$FILE" | tr -d '12' | awk '{print substr($0, 2, 1)}' | tr '\n' ' '); printf "$line $c\n"; done
%
Moreover, 23 distinct patterns were introduced to indicate hyphenation points
between vowels or between \verb|c|'s, \verb|r|'s, and \verb|s|'s.

%
% FILE="../data/default.TeX.pt-br.patterns"; grep -v "^1..$" "$FILE" | grep -v "2" | grep "3" | wc -l
%
Additionally, 8 patterns adhere to the \verb|1[gq]u4V|, signaling beneficial
hyphenation points before \verb|g| or \verb|q|, followed by a sequence of
\verb|u| and a vowel, with an inhibiting point between the \verb|u| and the
subsequent vowel.

Lastly, there is a single pattern represented as \verb|1-|, denoting that a
hyphen indeed serves as a beneficial hyphenation point."

% - contabilizar a quantidade de palavras que não são contempladas pela proposta do autor
% - analisar a necessidade de excluir alguma regra (ao invés de apenas acrescentar novas)

\section{Methodology}\label{sec-methodology}
To assess the performance of each set of rules, we require a list of correctly
hyphenated words in Portuguese. 
We utilized word frequency data to avoid selecting excessively rare words, creating a minimal representative set. 
A comprehensive list was compiled using data from online dictionaries and
Portuguese corpora. Additionally, we referred to Portuguese grammars to
validate hyphenation (\emph{translineation}) rules and manually corrected the data as
needed.

\subsection{Data collection}\label{sec-data-coll}

Initially, we selected
\href{https://www.linguateca.pt/cetenfolha/index_info.html}{CETENFolha}
CETENFolha as the source corpus, but this approach presented two issues: many
words were still missing, and the corpus was based on text from 1994, prior to
the implementation of the Orthographic Agreement.  The Orthographic Agreement
of the Portuguese Language was conducted in 1990, and its transition period
started in 2009, becoming mandatory in 2016. However, even after the agreement,
some words continue to have different spellings in the participating countries.
To address these idiosyncrasies, we aim to develop hyphenation rules that
accommodate variations in spelling. For instance, the word \emph{reception}
might be written as \emph{receção} (Portugal) or \emph{recepção} (Brazil); the
word \emph{action} might be written as \emph{acção} (Portugal) or \emph{ação}
(Brazil); and the word \emph{project} might be written as \emph{projecto}
(Portugal) or \emph{projeto} (Brazil).

% We have used the Portuguese Wikipedia dump as our Portuguese corpus. Texts in Wikipedia might be written by Brazilian or Portuguese contributors, the both spellings may appear.

% cat /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | tr -dc '0-9\n' | paste -sd+ | bc
% 302226482
% cat /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | awk '{CUMCOUNT+=$1; if(CUMCOUNT < 287115158) print;}' >/tmp/ptwikitop95.txt
% cat /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | awk '{CUMCOUNT+=$1; if(CUMCOUNT < 299204217) print;}' >/tmp/ptwikitop99.txt
% cat /tmp/ptwikitop99.txt | tr -d "0-9 " |  while read -r word; do if ! grep -q "^${word}," hyphenations.csv; then echo $word; fi; done > /tmp/ptwikitop99_notinlist.txt

Because of these idiosyncrasies, we decided to incorporate the list of words from
\href{https://www.palavras.net/}{Palavras NET} and afterwards we also included
words from the \href{https://pt.wikipedia.org}{Portuguese Wikipedia} dump.
Although the list of words created was huge, we kept only those words we could
find their hyphenation on at least one of the following online dictionaries: 
\href{https://michaelis.uol.com.br/}{Michaelis},
\href{https://dicionario.priberam.org/}{Priberam},
\href{https://pt.wiktionary.org}{Wikcionário},
\href{https://aulete.com.br/}{Aulete},
\href{http://www.portaldalinguaportuguesa.org/}{Portal da Língua Portuguesa} e
\href{https://www.dicio.com.br/}{Dicio}.
This leads to the creation of a dictionary with \DictionarySize{} words.

Table \ref{tblstats} presents the number of words with a given number of
hyphenation found in the dictionaries. It also presents the number of words
that have the highest agreed hyphenation. For example, the word \emph{como}
was hyphenated by the six dictionaries as \emph{co-mo}. It will add one to the
counts on both lines where the hyphenations number is 6 (first column). The
word \emph{sua} was hyphenated as \emph{su-a} in 4 dictionaries, \emph{sua} in
one dictionary and one dictionary did not provide its hyphenation. Therefore,
it will add one in the first line, second column (since hyphenations were found
in 5 dictionaries). It will also add one on the second line, third column
(since the most frequent hyphenation was found in 4 dictionaries). It is clear
now why we might have the number of same hyphenations (specially for small
values) greater the number of hyphenations found.

\begin{table}
\centering
\caption{Considering the six dictionaries used, the first line of this table
presents the number words that a given number of hyphenations were found in the
dictionaries. The second line presents the number of words that have a given
number of hyphenation as its most frequent form found in the
dictionaries.}\label{tblstats}
\small
\begin{tabular}{*{8}{l}}
    hyphenations & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
    \hline
    hyphenations found & \NumberOfSixHyphens{} & \NumberOfFiveHyphens{} &
    \NumberOfFourHyphen{} & \NumberOfThreeHyphens{} & \NumberOfTwoHyphens{} &
    \NumberOfOneHyphens{} & \NumberOfNoHyphens{} \\

    same hyphenations & \NumberOfSixAgrees{} & \NumberOfFiveAgrees{} &
    \NumberOfFourAgrees{} & \NumberOfThreeAgrees{} & \NumberOfTwoAgrees{} &
    \NumberOfOneAgrees{} & n/a
\end{tabular}
\end{table}

In summary, data from \emph{CETENFolha} and \emph{Palavras NET} were compiled.
Their words were hyphenated based on the hyphenation original \TeX{} rules and
contrasted with the dictionary results. The performance was appraised and the
errors systematized, in order to find in the literature possible rules that
could complement and improve it.

\section{Writing systems}\label{sec-writing-systems}

There is a wide range of writing systems. They might be categorized into
logographic, syllabic, and alphabetic writing systems.  Although they are
distinct categories, writing systems might be built on an interplay of these
categories \cite{coulmas2003}. The principles that guide an alphabetic writing
system or a phonemic writing system can vary significantly based on the
specific language and its history. Nevertheless, we might highlight here some
of its key points:
\begin{enumerate}[label=\arabic*)]
    \item the most usual principle consists on representing the sounds of a
	language with written symbols, using one or a combination of symbols to
	represent each sound (e.g. the orthographic system of Portuguese uses
	an alphabet of 27 letters that aim to represent different sounds of the
	language); 
    \item etymology also plays an important role, where the
	spelling of a word reflects its origins and history (e.g. the French
	language often reflects the Latin roots of words in its spelling);
    \item morphology is used as a guide in structuring many languages, where
	they use specific letters or symbols to indicate word endings,
	prefixes, or suffixes (e.g. Russian uses different forms of the
	Cyrillic alphabet to indicate gender and case in its nouns); 
    \item the way a language has evolved over time also contribute to its 
	written form (e.g. the spelling of many English words have changed 
	over time to reflect changes in pronunciation);
\end{enumerate}
The principles that guide orthographic systems for languages with logographic
writing systems like Chinese ideograms, cuneiform writing, and Egyptian
hieroglyphs differ significantly from those of alphabetic writing systems. In
logographic writing systems, individual symbols or characters represent entire
words or ideas rather than phonemes or sounds. As a result, the principles that
guide their orthographic systems are based more on semantic and visual
principles rather than phonemic principles.

\section{Grammar rules for hyphenation in Portuguese}\label{sec-grammar-port}

Portuguese uses an alphabetical writing system, ensuing that its spelling is
guided by phonological principles \cite{cagliari2015}. The fact that spelling
reflects, to some extent, the way words are pronounced, means that the
hyphenation of words is also influenced by the phonemic system, where words are
divided into syllables. It is noteworthy that different languages divide words
based on different principles, for example, English tends to be guided by
morphological principles, such as \emph{walk-ing}, \emph{un-happy},
\emph{work-s}, and \emph{ear-ly}. Also, other factors affect word hyphenation in
English, such as the presence of doubled consonants and digraphs\footnote{One
should not confuse consonant clusters with digraphs, since the latter represent
only one sound, while the former represent two or more sounds.}. There are also
a myriad of exceptions that might be grasped into rules.
% https://english.stackexchange.com/questions/385/what-are-the-rules-for-splitting-words-at-the-end-of-a-line

Merely stating that an orthographic system is guided by phonological issues
does not necessarily mean that its hyphenation rules directly reflect the way a
word is pronounced -- this is because Portuguese does not have a direct
letter-sound correspondence, and vice versa.  The orthographic system operates
according to its own specific rules. For instance, although consonant clusters that
produce a single sound (i.e., digraphs) are pronounced within a single syllable
in words like \emph{chave}, \emph{ilha}, \emph{shampoo}, \emph{carro}, and
\emph{pássaro}, their representation in writing determines how they are
separated.  Specifically, different consonants within a digraph must remain
together, while identical consonants are separated. As a result, we have
\emph{cha-ve}, \emph{i-lha}, and \emph{sham-poo}, but \emph{car-ro}.

%When there is no space left in line for a whole word, it might be split in two
%using a hyphen as an indication of such procedure.

In Portuguese, hyphenations are allowed on syllables boundaries and, in
general, follow phonological principles.  According to the Grammar
\cite{cunha2016,bergstrom2011,cegalla2008}, some rules might still apply:
\begin{enumerate}
    \item\label{rule-di-triphthong} diphthong or triphthong should not be split
	(e.g. \emph{m\underline{\emph{ui}}-to}, \emph{Pa-ra-g\underline{\emph{uai}}});
    \item\label{rule-unstressed} the sequences \emph{ia}, \emph{ie}, \emph{io}, \emph{oa}, \emph{ua},
	\emph{ue} and \emph{uo}, when in final unstressed position, should not be split 
	(e.g. \emph{gló-r\underline{\emph{ia}}}, \emph{vi-tó-r\underline{\emph{ia}}}, 
	\emph{cá-r\underline{\emph{ie}}}, \emph{es-pé-c\underline{\emph{ie}}}, 
	\emph{Má-r\underline{\emph{io}}}, \emph{má-g\underline{\emph{oa}}}, 
	\emph{ré-g\underline{\emph{ua}}}, \emph{tê-n\underline{\emph{ue}}}, 
	\emph{con-tí-g\underline{\emph{uo}}}, \emph{am-bí-g\underline{\emph{uo}}});
    \item\label{rule-c-clusters} consonant clusters starting a syllable should not be
	split (e.g. \emph{\underline{\emph{pn}}eu-má-ti-co}, \emph{\underline{\emph{ps}}i-có-lo-go}, 
	\emph{\underline{\emph{mn}}e-mô-ni-co});
    \item\label{rules-digraphs-ns} the digraphs	\emph{ch}, \emph{lh}, \emph{nh}
	should not be split (e.g. \emph{ra-\underline{\emph{ch}}ar},
	\emph{a-bro-\underline{\emph{lh}}os}, \emph{ma-\underline{\emph{nh}}ã};
    \item\label{rule-guqu} bigrams like \emph{gu} and \emph{qu} whose vowel \emph{u} is not 
	pronounced are never separated from the vowel or diphthong that follows it 
	(e.g. \emph{ne-\underline{\emph{gu}e}}, \emph{pe-\underline{\emph{qu}e}});
    \item\label{rule-nasalization} since they are digraphs, a vowel and its following 
	nasalization marker (a graphic nasal consonant) should not be split (e.g. \emph{\underline{\emph{am}}-bição},
	\emph{m\underline{\emph{an}}-cha});
    \item\label{rule-decreasing} decreasing diphthongs should not be split (e.g. 
	\emph{\underline{\emph{ai}}-ro-so}, \emph{ca-d\underline{\emph{ei}}-ra}, 
	\emph{o-ra-ç\underline{\emph{ão}}});
    \item\label{rule-rising} rising diphthong should not be split (e.g.
	\emph{a-b\underline{\emph{ai}}-xo}; \emph{c\underline{\emph{au}}-te-la}, 
	\emph{pa-p\underline{\emph{éi}}s}, \emph{cha-p\underline{\emph{éu}}}, 
	\emph{pre-f\underline{\emph{ei}}-to}, \emph{r\underline{\emph{eu}}-nir},
	\emph{n\underline{\emph{oi}}-te}; \emph{ca-la-b\underline{\emph{ou}}-ço}, 
	\emph{as-te-r\underline{\emph{ói}}-de}; 
	\emph{re-tri-b\underline{\emph{ui}}});
    \item\label{rule-singlev} disyllables whose syllable has a single vowel should 
	not be split (e.g. \emph{\underline{\emph{a}}to}, \emph{ru\underline{\emph{a}}}, 
	\emph{\underline{\emph{ó}}dio}, \emph{\underline{\emph{u}}nha});
    \item\label{rule-orphan} words with more than two syllables, when divided, cannot 
	isolate a syllable composed of a single vowel (e.g. \emph{\underline{\emph{a}}gos-to}, 
	\emph{la-go\underline{\emph{a}}}, \emph{\underline{\emph{i}}da-de});

    \hspace{10em} \hbox to 5cm{\leaders\hbox to 10pt{\hss . \hss}\hfil}

    \item\label{rule-hiatus} hiatus vowels and those vowel sequences where each vowel
	belongs to a different syllable should be split (e.g.
	\emph{sa\underline{\emph{-ú-}}de}, \emph{ra\underline{\emph{-i-}}nha}, \emph{d\underline{\emph{o-e}}r},
	\emph{v\underline{\emph{o-o}}s}), the same procedure is used splitting diphthongs
	in different syllables (e.g. \emph{c\underline{\emph{ai-ai}}s}) or diphthong and
	vowel in different syllables (e.g. \emph{en-s\underline{\emph{ai-o}}s});
    \item\label{rule-consonants} consonant sequences, when in different syllables, should
	be split (e.g. \emph{a\underline{\emph{f-t}}a}, \emph{a\underline{\emph{b-d}}i-car},
	\emph{re\underline{\emph{s-c}}i-são}, \emph{a\underline{\emph{b-s}}o-lu-to});
    \item\label{rule-digraphs} the following consonant digraphs should be split:
	\emph{rr}, \emph{ss}, \emph{mm}, \emph{nn}, \emph{sc}, \emph{sç} and
	\emph{xc} (e.g. \emph{te\underline{\emph{r-r}}a}, \emph{pro-fe\underline{\emph{s-s}}or},
	\emph{co-mu\underline{\emph{m-m}}en-te}, \emph{co\underline{\emph{n-n}}os-co},
	\emph{de\underline{\emph{s-c}}er}, \emph{cre\underline{\emph{s-ç}}a}, \emph{e\underline{\emph{x-c}}e-der}).
\end{enumerate}

Note that rules \ref{rule-singlev} and \ref{rule-orphan} are fiddling rules in \TeX{} hyphenation, 
since \TeX{} already has directives to avoid hyphenated widows and orphans. 
As mentioned in \Cref{sec-intro}, the variables \verb|lefthyphenmin| and \verb|righthyphenmin|
are used to control the minimum length for fragments of hyphenated words.
This is the ground why we will not consider rules \ref{rule-singlev} and \ref{rule-orphan}
when analysing and elaborating the hyphenation rules.

In some situations a hyphen should be repeated in the start of the following
line: 
\begin{enumerate}
\item when a compound word using a hyphen is split across lines (e.g.
    \emph{couve-/-flor}, \emph{ex-/-presidente});
\item when a different meaning could be conveyed when splitting a pronoun (e.g.
    \emph{prazer de ver-/-me}\footnote{Possible conveyed meanings: 
    \emph{pleasure in seeing me} or \emph{worm's pleasure}.}). 
\end{enumerate}

It should be avoided to leave a single vowel in the end or start of a line (e.g.
the word \emph{san-guí-nea} should be split only on the hyphen gave in the
example, avoiding the split that would leave the final vowel \emph{a}
left alone in the start of a new line: \emph{sanguíne-a}). It is also considered a better approach
to avoid splitting disyllables made of four letters (e.g. \emph{para},
\emph{como}, \emph{cede}).

Systematizing the rules that guide syllable boundaries and, consequently,
hyphenation in Portuguese proved to be fundamental for the improvement of the
\TeX{} hyphenator rules, since they indicate the possible or prohibited contexts
of hyphenation. Expomos, a seguir, uma análise comparativa...

... to continue
\vspace{10ex}

\section{Creating the dictionary}
A \emph{bash} script was created to crawl on six online dictionaries and extract a word hyphenation, when found.

\section{Updating the rules}
Let us remember that the standard \TeX{} hyphenation rules, proposed by
\cite{rezende1987,hyphpt}, when tested against a set of \num{\DictionarySize}
words, present 1,368 (????) errors, of which 1,273 (????) correspond to places
where the hyphenation was not carried out and 125 (????) to erroneous markings.
Faced with the errors made by the standard hyphenator, a set of \num{\NumberOfNewRules} rules were
established, which reduced the errors to 132 (???), with 118 (????) places
where the hyphenation was not marked and 35 (????) where there was erroneous
marking. Such rules are presented and exemplified below, as well as the
necessary exception rules and an example in which the application of the rule
is demonstrated.

\subsection{Results from the new set of rules}
Although the XXX erroneous words have some common characteristics, which would
allow, to some extent, to reduce the amount of hyphenation errors, there is a
limitation arising from the way \TeX{}'s rules are conceived. We present,
below, the systematists found in these data that could be predicted by another
structure of rules:

\begin{description} 
    \item[Morphological determination:] Prefixes such as \emph{re-},
	\emph{sub-}, \emph{cyber-}, and \emph{auto-} require separating the
	prefix from its stem, what runs into phonological issues. For example,
	\emph{reiniciar} and \emph{sublinhar}, as they contain prefixes, should
	be hyphenated as \emph{re-i-ni-ci-ar} and \emph{sub-li-nhar},
	respecting the morphological formation of the word. However,
	considering that the Portuguese language hyphenates its words from
	syllabic phonological correlates, words that require morphological
	information, such as those, do not have their hyphenation performed
	properly. They add up, in our corpus, to twelve words (eight words with
	\emph{re-} and one case with \emph{auto-}, \emph{meta-}, \emph{sub-},
	\emph{cyber-}, \emph{geo-} or \emph{neuro-}).

    \item[Foreignness:] There is a group of 20 words in the corpus that are
	terminologies or words that were incorporated into the Portuguese
	language without full phonological adaptation, such as
	\emph{darwinismo}, \emph{quilowatt}, and \emph{esfiha}. The lack of
	adaptation makes the phonological pattern very specific to the word,
	making it impossible to incorporate in \TeX{}'s rules alongside the
	other rules. The solution is to add them to the exception word list.

    \item[Word-initial consonant clusters:] Portuguese has few cases of
	consonant clusters at the beginning of a word. They, in general, are
	etymological remnants and are currently unproductive in the language,
	since there are no neologisms with this pattern. Encounters like
	\emph{ps-} and \emph{pn-} are more frequent, as they are present in
	words like \emph{psicologia} and \emph{pneu}, which have an average
	frequency in the language. They might be predicted by rules, which
	comprise XXX words. However, there are consonant clusters that are
	found in very specific and low frequency words. Although possible, it
	is not worth adding very specific rules using the consonant clusters
	contained in the words \emph{dzeta}, \emph{gnu}, \emph{cnidário},
	\emph{ftálico}, and \emph{gnaisse} -- that is, just five words.

    \item[Abbreviations, Acronyms, or Initialisms:] Whether for efficiency, 
	convenience, clarity or jargon, it is usual the usage of a shortening 
	version of a word or a phrase. 
	Abbreviation a way to shorten a word of phrase. Some examples in our
	Portuguese corpus are: 
	\emph{etc.}\footnote{Latin expression \emph{et cetera}, meaning 
	`and other similar things'.}, 
	\emph{Dr.}\footnote{\emph{doutor} (doctor, person with PhD title, but popularly used to designate an erudite individual)}, 
	\emph{Exmo.}\footnote{\emph{Excelentíssimo} (honourable)},
	\emph{cap.}\footnote{\emph{capítulo} (chapter)}, \emph{Univ.}\footnote{Universidade (university)}, 
	\emph{ed.}\footnote{\emph{edição} (edition)}, 
	\emph{s.n.}\footnote{sine nomine, Latin expression meaning `without a name', 
	mostly used in the context of publishing.}.
	Initialism consists of a set of initial 
	letters of words that form an shortened version. They are not always 
	understood by the set of hyphenation rules described in this work, 
	since they do not follow orthographic (or phonotactic) standards of 
	the language. In the corpus, it is possible to find some abbreviations, 
	such as \emph{SESC}\footnote{Serviço Social do Comércio}, 
	\emph{INSS}\footnote{\emph{Instituto Nacional do Seguro Social} (National Institute of Social Security)}, 
	\emph{PCdoB}\footnote{\emph{Partido Comunista do Brasil} (Communist Party of Brazil)},
	\emph{PM}\footnote{\emph{Polícia Militar} (military police)}, and 
	\emph{UFRJ}\footnote{Universidade Federal do Rio de Janeiro}.
	Acronym is a type of abbreviation that shortens a phrase by combining
	the first letter (or letters) of each word, creating a new pronounceable word.
	Some example from the corpus are: 
	\emph{Anatel}\footnote{\emph{Agência Nacional de Telecomunicações} (National Telecommunications Agency)},
	\emph{Ovni}\footnote{\emph{Objeto voador não identificado} (unidentified flying object - UFO)},
	\emph{Sida}\footnote{\emph{Síndrome da Imunodeficiência Adquirida} (acquired immunodeficiency syndrome - AIDS)} (in Portugal) and
	\emph{Mercosul}\footnote{\emph{Mercado Comum do Sul} (Southern Common Market).}.
\end{description}

\section{Creating a new set of rules using Patgen}

\section{Comparing the set of rules} 
Let us remember that the standard \TeX{} hyphenation rules, proposed by
\cite{hyphpt}, amount to XXX errors, in a set of XXXX words.  Among those, XXX
correspond to places where the hyphenation was not carried out and XXX to
mismarkings. Faced with the errors made by the standard hyphenator, a set of
\NumberOfNewRules{} rules were established, which reduced the errors to XXXX,
with XXXX unmarked hyphenation places and XXX mismarkings.  Such rules are
presented and exemplified below, as well as the necessary exception rules and
an example in which the application of the rule is demonstrated.

\section{Old Text}

Foi criado um conjunto de \NumberOfNewRules{} regras para aprimorar as hifenizações no
português. Analisando um conjunto de 37798 palavras, as regras padrão apresentam
1368 erros (1273 pontos de hifenização não marcados e 125 pontos marcados
erroneamente)\footnote{A soma não é 1368 pois temos 30 casos em que os dois
tipos de erros ocorrem.}, enquanto as novas regras apresentaram 132 erros (118
pontos de hifenização não marcados e 35 pontos marcados
erroneamente)\footnote{35 palavras apresentam ambos tipos de erro.}.
%%% <<< fazer correção após correção manual das hifenizações tidas como corretas.

Foi criado um conjunto de 37798 palavras obtidas do corpus
\href{https://www.linguateca.pt/cetenfolha/index_info.html}{CETENFolha},
juntamente com palavras do \href{https://www.palavras.net/}{Palavras NET} e
algumas palavras do corpus baseado na
\href{https://pt.wikipedia.org}{Wikipedia}. 
Foram mantidas apenas as palavras
para as quais encontrou-se a hifenização nos seguintes dicionários online:
\href{https://michaelis.uol.com.br/}{Michaelis},
\href{https://dicionario.priberam.org/}{Priberam},
\href{https://pt.wiktionary.org}{Wikcionário},
\href{https://aulete.com.br/}{Aulete},
\href{http://www.portaldalinguaportuguesa.org/}{Portal da Língua Portuguesa} e
\href{https://www.dicio.com.br/}{Dicio}.
Adotamos como hifenização `correta' aquela que ao menos 4 dicionários
concordavam. Algumas hifenizações foram realizadas manualmente pelo autor.
Considerando a frequência de ocorrência das palavras nos corpus, o conjuto de
palavras escolhido representa 55\% do total das ocorrências relativo ao
Wikipedia e 58\% em relação ao CETENFolha.
% Corpus Wikipedia
%
% 166341424 / 302226482 = 0.55
% 
% soma da freq. de occ. das palavras utilizadas
% cat portuguese_4_utf8.dic portuguese_extra_utf8.dic | tr -d '-' | while read -r word; do grep "[0-9]\+ ${word}$" /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | head -n 1 | tr -dc '0-9\n'; done | paste -s -d+ - | bc 
% = 166341424
%
% soma da freq. de occ. de todas palavras no corpus
% cat /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | tr -dc '0-9\n' | paste -s -d+ - | bc
% = 302226482
%
% CETENFolha
% 13314084 / 22728804 = 0.5857
Este percentual é relativamente baixo, mas deve-se ao fato de que não obtivemos
sucesso em encontrar a hifenização de muitas palavras nos dicionários
analisados e também por não observarmos um consenso de ao menos 4 dicionários
com uma mesma hifenização.

Analisando os erros resultantes da hifenização utilizando as regras padrão do
TeX, foram estabelecidas algumas regras como forma de abarcar um conjunto maior
de palavras. Tais regras são elencadas a seguir, expondo também as regras de
exceção necessárias e um exemplo onde a regra será utilizada.

\begin{enumerate}
\item\label{rulegrp_psi} regra: \texttt{p2si}, \texttt{p2sí} -- psicologia, psíquico\\
exceção: \texttt{p3sia}, \texttt{p3sin} -- epilepsia, rodopsina

\item\label{rulegrp_tc} regra: \texttt{t2c} -- tchau%, tcheco

\item\label{rulegrp_pneu} regra: \texttt{p2neu} -- pneumonia

\item\label{rulegrp_gno} regra: \texttt{.g2no}, \texttt{.g2nó} -- gnomo, gnóstico

\item\label{rulegrp_pt} regra: \texttt{.p2t} -- ptose%, pterossauro

\item\label{rulegrp_cza} regra: \texttt{c2za} -- czar

\item\label{rulegrp_s2} regra: \texttt{.s2} -- stalinismo

\item\label{rulegrp_t2} regra: \texttt{.t2} -- tsunami

\item\label{rulegrp_mn} regra: \texttt{.m2n} -- mnemônico

\item\label{rulegrp_zz} regra: \texttt{z1z}, \texttt{p1p}, \texttt{z2z.} -- pizza, shopping, jazz

\item\label{rulegrp_sub} regra: \texttt{su2b3r}, \texttt{su2b3l} -- sublunar, subrotina
exceção: \texttt{su3b4li1nh}, \texttt{su3b4li1ma}, \texttt{su3b4li1me}, \texttt{su3b4li1mid} -- sublinhar, sublimar, sublime, sublimidade

\item\label{rulegrp_so} regra: \texttt{1sô} -- subsônico

\item\label{rulegrp_pseu} regra: \texttt{1p2seu1d} -- pseudônimo

\item\label{rulegrp_ein} regra: \texttt{e1in1c}, \texttt{e1in1f}, \texttt{e1in1g}, \texttt{e1in1s}, \texttt{e1in1t}, \texttt{e1in1v} -- reincidência, reinfecção, reingressa, reinserção, reintegração, reinventar
exceção: \texttt{e1in2st}, \texttt{e1in2sc} -- reinstalado, reinscrever

\item\label{rulegrp_oin} regra: \texttt{o1in} -- coincidência, agroindustrial, bioinformática, amendoins, gastrointestinal
exceção: \texttt{o1ina} -- boina

\item\label{rulegrp_eim} regra: \texttt{e1im} -- reimpresso, coimbra, coimperador
exceção: \texttt{que2i}, \texttt{te2im} -- queimar, queimadura, teimosia, tirateima

\item\label{rulegrp_aa} regra: \texttt{a1â}, \texttt{a1ã}, \texttt{a1é}, \texttt{a1í}, \texttt{a1ó}, \texttt{a1ô}, \texttt{a1ú}, \texttt{e1á}, \texttt{e1â}, \texttt{e1ã}, \texttt{e1ê}, \texttt{e1í}, \texttt{e1ó}, \texttt{e1ô}, \texttt{e1ú}, \texttt{i1á}, \texttt{i1ã}, \texttt{i1ã}, \texttt{i1é}, \texttt{i1ó}, \texttt{i1ú}, \texttt{o1á}, \texttt{o1ã}, \texttt{o1é}, \texttt{o1ê}, \texttt{o1í}, \texttt{o1ó}, \texttt{u1á}, \texttt{u1ã}, \texttt{u1é}, \texttt{u1ê}, \texttt{u1í} --  abraâmico, abraão, aéreo, país, caótico, faraônico, saúde, balneário, oceânico, campeã, veêm, veículo, teórico, napoleônico, conteúdo, diário, região, soviético, periódico, viúva, razoável, joão, poético, boêmia, heroísmo, alcoólico, usuário, itapuã, suécia, cauê, suíça \\
exceção: \texttt{1gu2é}, \texttt{1gu2ê}, \texttt{1gu2í}, \texttt{1qu2á}, \texttt{1qu2é}, \texttt{1qu2ê}, \texttt{1qu2í} -- alguém, português, linguística, aquático, inquérito, sequência, química

\item\label{rulegrp_air} regra: \texttt{a1ir.}, \texttt{u1ir.} -- sair, diminuir

\item\label{rulegrp_ia} regra: \texttt{í1a} -- baía

\item\label{rulegrp_ain} regra: \texttt{a1ind}, \texttt{a1i1nh} -- ainda, rainha

\item\label{rulegrp_oin} regra: \texttt{o1i1nh} -- moinho

\item\label{rulegrp_aui} regra: \texttt{au1i1c}, \texttt{du1i1c}, \texttt{u1i1ç}, \texttt{u1i1d}, \texttt{cu2i}, \texttt{dru2i}, \texttt{flu2id}, \texttt{bu1i1n}, \texttt{cu1i1n}, \texttt{fu1i1n}, \texttt{nu1i1n}, \texttt{ru1i1na}, \texttt{ru1i1no}, \texttt{su1i1ti}, \texttt{tu1i1ti}, \texttt{u1iz} -- cacauicultor, sanduicheira, constituição, continuidade, cuidador, druida, fluido, tabuinha, picuinha, fuinha, genuinamente, arruinado, arruinou, jesuitismo, intuitivo, juizado 
exceção: \texttt{cu3i1da1de}, \texttt{bu1i2n1d}, \texttt{bu1i2n1t}  -- acuidade, contribuindo, contribuinte

\item\label{rulegrp_gua} regra: \texttt{1gu4á}, \texttt{1gu4ã}, \texttt{1qu4ã} -- jaraguá, saguão, quão

\item\label{rulegrp_iur} regra: \texttt{i1ur} -- diurno

\item\label{rulegrp_io} regra: \texttt{í1o} -- íon

\item\label{rulegrp_proi} regra: \texttt{pro1i1b} -- proibição

\item\label{rulegrp_iun} regra: \texttt{i1un} -- triunfar

\item\label{rulegrp_quo} regra: \texttt{1qu4ó}, \texttt{1qu4â} -- quórum, equânime

%\item\label{rulegrp_uo} regra: \texttt{u3ó} -- fluór

\item\label{rulegrp_eo} regra: \texttt{é3o} -- alvéolo

\item\label{rulegrp_ia} regra: \texttt{i4a.}, \texttt{i4e.}, \texttt{i4o.}, \texttt{o4a.}, \texttt{u4a.} -- economia, espécie, vazio, destoa, institua

\item\label{rulegrp_lo} regra: \texttt{1lô} -- camelô

\item\label{rulegrp_co} regra: \texttt{1cô} -- recôncavo

\item\label{rulegrp_bo} regra: \texttt{1bô}, \texttt{1dô}, \texttt{1fô}, \texttt{1gô}, \texttt{1pô}, \texttt{1mô}, \texttt{1nô}, \texttt{1rô}, \texttt{1tô}, \texttt{1vô}, \texttt{1xô}, \texttt{1zô} -- robô, judô,  telefônica, xangô, capô, sumô, econômico, tarô, chatô, vovô, saxônia, amazônia

\item\label{rulegrp_cco} regra: \texttt{1çô} -- maçônico

% As sequências gráficas ia, ie, io, oa, ua quando em posição
% final átona, que são normalmente pronunciadas como ditongos
% crescentes, mas que podem corresponder foneticamente a duas
% vogais (hiato), sobretudo quando pronunciadas pausadamente:
% vitó-ria, espé-cie, exercí-cio, nó-doa, lé-gua, tê-nue, ambí-guo.
    \setcounter{numberRulesGroups}{\value{enumi}}
\end{enumerate}

The \NumberOfNewRules{} rules were grouped above in a list of
\arabic{numberRulesGroups} types of rules. They may be further organized into four
large groups. The first, which comprises rules \ref{rulegrp_psi} to
\ref{rulegrp_zz}, includes consonant clusters such as \emph{czar}, \emph{ptose}
and \emph{gnomo}. They, unlike the examples that will be exposed in section
\ref{secxxx} (exceptions - item 9 currently), present a set of derived words,
which makes their marking advantageous in view of the number of cases that are
included in this marking rule. The second group, comprising rules
\ref{rulegrp_sub} to \ref{rulegrp_eim}, delimits the morphological boundary
between prefixes and radicals. As noted, although phonological issues guide the
separation of numerous words in Portuguese, there are also those that are
guided by morphology. This is the case of words that have the prefixes
\emph{sub-} and \emph{re-}, such as \emph{sublunar} and \emph{reinserção}. The
third group, comprising rules \ref{rulegrp_aa} to \ref{rulegrp_ia}, seeks to
understand a set of words that have vowel combinations that do not follow the
general rules. This is because the Portuguese language has vowel encounters
with the second vowel graphically marked that can be separated, forming
hiatuses, such as \emph{caótico}, \emph{balneário} and \emph{razoável}, while
there are also words with a similar structure that constitute a diphthong, such
as português, \emph{alguém} and \emph{linguística}. It is remarkable, of
course, that the latter are formed by the digraphs \emph{qu-} and \emph{gu-},
while the former by vowels other than \emph{i} and \emph{u}. The fourth and
last group, in turn, which comprises rules \ref{rulegrp_lo} to
\ref{rulegrp_cco}, which are counterparts of rules that were already in the
default rules, but did not contemplate the cases with certain accents. They
were then added to encompass words such as \emph{camelô}, \emph{recôncavo},
\emph{amazônia}, and \emph{maçônico}.

It is important to highlight that the words included in these rules are, in general, of low frequency and were incorporated into the Portuguese language without a phonotactic adaptation, which causes these idiosyncrasies and exceptions to the language. They are characterized as marked cases, since it is not possible to defend that they reproduce a phonological pattern of the language because they are not productive, that is, they are not taken as an example and/or derive new words.

O conjunto de \NumberOfNewRules{} regras é apresentado a seguir:
\begin{multicols}{6}
    \footnotesize
    \verbatiminput{../data/patch.TeX.pt-br.patterns}
\end{multicols}

A Tabela \ref{tab-resultados} apresenta os erros cometidos no conjunto das
37798 palavras analisadas. As palavras omitidas são aquelas que foram
hifenizadas corretamente pelos dois conjuntos de regra. Utilizamos aqui a
marcação adotada pelo \emph{Patgen}: \texttt{*} para indicar pontos de
hifenização corretamente marcados, \texttt{.} para indicar pontos de hifenização
marcados erroneamente e \texttt{-} para indicar pontos de hifenização não
encontrados. Os pontos marcados erroneamente e aqueles não encontrados são
considerados erros. Quando ocorre algum erro de hifenização, marcaremos a
palavra com \xmark. Quando não houver erro algum, marcaremos a palavra com
\cmark. Ao final, podemos verificar que as regras padrão apresentaram 1368 erros
(3,62\%) e as novas regras 132 erros (0,35\%).

\begin{longtable}{l l l}
\caption{Comparativo dos erros cometidos pelos dois conjuntos de regras de
hifenização.}\label{tab-resultados}\\
    \hline
    palavra & regras padrão & regras adicionais \\
    \hline
    \endhead
    \input{comparedresults.tex}
\end{longtable}

\printbibliography

\end{document}