### 2024-06-14 article/article.tex

% xelatex --shell-escape article.tex 
% pandoc -s article.tex --pdf-engine=xelatex --pdf-engine-opt=-shell-escape -o article.html
\documentclass{article}
\usepackage[a4paper,margin=2cm,nohead,foot=1cm]{geometry}
\usepackage[english]{babel}
%\usepackage{testhyphens}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{amssymb}% http://ctan.org/pkg/amssymb
\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\ding{51}}%
\newcommand{\xmark}{\ding{55}}%
\usepackage{longtable}
\usepackage[style=authoryear, backend=biber]{biblatex}
\addbibresource{references.bib}
\usepackage[inline]{enumitem}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{fancyvrb}
\usepackage{soul}
\usepackage{siunitx}
\sisetup{
  group-separator = {\,},
  group-minimum-digits = 3,
  output-decimal-marker = {.},
}
\usepackage{booktabs}

%\usepackage[textsize=tiny,textwidth=3.5cm,color=blue!10]{todonotes}

%%%%%%% https://tex.stackexchange.com/questions/16790/write18-capturing-shell-script-output-as-command-variable
\usepackage{xparse}
\ExplSyntaxOn
\NewDocumentCommand{\captureshell}{som}
 {
  \sdaau_captureshell:Ne \l__sdaau_captureshell_out_tl { #3 }
  \IfBooleanT { #1 }
   {% we may need to stringify the result
    \tl_set:Nx \l__sdaau_captureshell_out_tl
     { \tl_to_str:N \l__sdaau_captureshell_out_tl }
   }
  \IfNoValueTF { #2 }
   {
    \tl_use:N \l__sdaau_captureshell_out_tl
   }
   {
    \tl_set_eq:NN #2 \l__sdaau_captureshell_out_tl
   }
 }

\tl_new:N \l__sdaau_captureshell_out_tl

\cs_new_protected:Nn \sdaau_captureshell:Nn
 {
  \sys_get_shell:nnN { #2 } { } #1
  \tl_trim_spaces:N #1 % remove leading and trailing spaces
 }
\cs_generate_variant:Nn \sdaau_captureshell:Nn { Ne }
\ExplSyntaxOff
%%%%%%%

\usepackage{rotating}
\usepackage{bytefield}
%\usepackage{xstring}

% Define a new counter for the custom list of rules
\newcounter{numberRulesGroups}
% Define a new list environment for rules
\newlist{rules}{enumerate}{1}
\setlist[rules,1]{label=\arabic*, ref=\arabic*, before=\setcounter{numberRulesGroups}{0}}

% Custom label format for the rules environment
\crefname{rulesi}{rule}{rules}
\Crefname{rulesi}{Rule}{Rules}

% Custom format for multiple references
\crefformat{rulesi}{rule~#2#1#3}  % Use only the number for single reference
\crefrangeformat{rulesi}{#3#1#4 to #5#2#6}  % Use the range format for range of references
\crefmultiformat{rulesi}{rules~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3} % Use the number format for multiple references
\Crefmultiformat{rulesi}{Rules~#2#1#3}{ and~#2#1#3}{, #2#1#3}{ and~#2#1#3} % Use the number format for multiple references

% https://tex.stackexchange.com/questions/1230/reference-name-of-description-list-item-in-latex
\makeatletter
\let\orgdescriptionlabel\descriptionlabel
\renewcommand*{\descriptionlabel}[1]{%
  \let\orglabel\label
  \let\label\@gobble
  \phantomsection
  \edef\@currentlabel{#1\unskip}%
  %\edef\@currentlabelname{#1}%
  \let\label\orglabel
  \orgdescriptionlabel{#1}%
}
\makeatother

\title{Improving the portuguese hyphenation rules}
\author{Leonardo Araújo \and Aline de Lima Benevides}
\begin{document}
\VerbatimFootnotes % Enable fancyvrb package to work in footnotes
\maketitle

\input{defvariables.tex}

\begin{abstract}
    Portuguese hyphenation rules are available for more than 35
    years and have done a good job. Nonetheless they still make mistakes and
    leave some hyphenation points unmarked. Although most undetected
    hyphenation points are located near word boundaries, what will be
    irrelevant for \TeX{} typographic purposes, they are still useful to
    hyphenate proper nouns, new words or pseudoword, and for usage in other
    applications, such as text-to-speech conversion. A list of
    \num{\DictionarySize} hyphenated words acquired from online dictionaries was
    used along \emph{patgen} to create improved rules, leading to better hyphenation
    of Portuguese words.
\end{abstract}

\input{introduction.tex}
\input{patterns.tex}
\input{patgen.tex}
\input{texrulespt.tex}
\input{methodology.tex}
\input{writingsystems.tex}
\input{portuguesespelling.tex}
\input{dictionary.tex}
\input{resultsdefaultrules.tex}
\input{updatingrules.tex}
\input{limitations.tex}
\input{patgenrules.tex}
\input{hyphenpatches.tex}
\input{conclusion.tex}

%\section{Comparing the set of rules} 
%Let us remember that the standard \TeX{} hyphenation rules, proposed by
%\textcite{hyphpt}, amount to XXX errors, in a set of XXXX words.  Among those, XXX
%correspond to places where the hyphenation was not carried out and XXX to
%mismarkings. Faced with the errors made by the standard hyphenator, a set of
%\num{\NumberOfNewRules} rules were established, which reduced the errors to XXXX,
%with XXXX unmarked hyphenation places and XXX mismarkings.  Such rules are
%presented and exemplified below, as well as the necessary exception rules and
%an example in which the application of the rule is demonstrated.

%\section{Universal syllabic pattern generation}
%Universal syllabic pattern generation by Jakub Máca, Petr Sojka, Ondřej Sojka (TUG 2023).

%% ==== OLD TEXT ====
%\section{Hyphenation rule patches}\label{sec-hyph-rpatches}
%OLD TEST
%
%OLD TEST
%
%Foi criado um conjunto de \num{\NumberOfNewRules} regras para aprimorar as hifenizações no
%português. Analisando um conjunto de \num{37798} palavras, as regras padrão apresentam
%1368 erros (1273 pontos de hifenização não marcados e 125 pontos marcados
%erroneamente)\footnote{A soma não é 1368 pois temos 30 casos em que os dois
%tipos de erros ocorrem.}, enquanto as novas regras apresentaram 132 erros (118
%pontos de hifenização não marcados e 35 pontos marcados
%erroneamente)\footnote{35 palavras apresentam ambos tipos de erro.}.
%%%% <<< fazer correção após correção manual das hifenizações tidas como corretas.
%
%Foi criado um conjunto de 37798 palavras obtidas do corpus
%\href{https://www.linguateca.pt/cetenfolha/index_info.html}{CETENFolha},
%juntamente com palavras do \href{https://www.palavras.net/}{Palavras NET} e
%algumas palavras do corpus baseado na
%\href{https://pt.wikipedia.org}{Wikipedia}. 
%Foram mantidas apenas as palavras
%para as quais encontrou-se a hifenização nos seguintes dicionários online:
%\href{https://michaelis.uol.com.br/}{Michaelis},
%\href{https://dicionario.priberam.org/}{Priberam},
%\href{https://pt.wiktionary.org}{Wikcionário},
%\href{https://aulete.com.br/}{Aulete},
%\href{http://www.portaldalinguaportuguesa.org/}{Portal da Língua Portuguesa} e
%\href{https://www.dicio.com.br/}{Dicio}.
%Adotamos como hifenização `correta' aquela que ao menos 4 dicionários
%concordavam. Algumas hifenizações foram realizadas manualmente pelo autor.
%Considerando a frequência de ocorrência das palavras nos corpus, o conjuto de
%palavras escolhido representa 55\% do total das ocorrências relativo ao
%Wikipedia e 58\% em relação ao CETENFolha.
%% Corpus Wikipedia
%%
%% 166341424 / 302226482 = 0.55
%% 
%% soma da freq. de occ. das palavras utilizadas
%% cat portuguese_4_utf8.dic portuguese_extra_utf8.dic | tr -d '-' | while read -r word; do grep "[0-9]\+ ${word}$" /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | head -n 1 | tr -dc '0-9\n'; done | paste -s -d+ - | bc 
%% = 166341424
%%
%% soma da freq. de occ. de todas palavras no corpus
%% cat /ms/downloads/samples/wikipedia/ptwiki-latest-pages-articles-multistream_wordlist.txt | tr -dc '0-9\n' | paste -s -d+ - | bc
%% = 302226482
%%
%% CETENFolha
%% 13314084 / 22728804 = 0.5857
%Este percentual é relativamente baixo, mas deve-se ao fato de que não obtivemos
%sucesso em encontrar a hifenização de muitas palavras nos dicionários
%analisados e também por não observarmos um consenso de ao menos 4 dicionários
%com uma mesma hifenização.
%
%Analisando os erros resultantes da hifenização utilizando as regras padrão do
%TeX, foram estabelecidas algumas regras como forma de abarcar um conjunto maior
%de palavras. Tais regras são elencadas a seguir, expondo também as regras de
%exceção necessárias e um exemplo onde a regra será utilizada.
%
%\begin{rules}
%\item\label{rulegrp_psi} regra: \texttt{.p2si}, \texttt{.p2sí} -- psicologia, psíquico % update (word-initial pattern, exception not necessary (?)) 
%%\item\label{rulegrp_psi} regra: \texttt{p2si}, \texttt{p2sí} -- psicologia, psíquico\\
%%exceção: \texttt{p3sia}, \texttt{p3sin} -- epilepsia, rodopsina
%
%\item\label{rulegrp_tc} regra: \texttt{t2c} -- tchau%, tcheco
%
%\item\label{rulegrp_pneu} regra: \texttt{p2neu} -- pneumonia
%    
%\item\label{rulegrp_gno} regra: \texttt{.g2no}, \texttt{.g2nó} -- gnomo, gnóstico
%
%\item\label{rulegrp_pt} regra: \texttt{.p2t} -- ptose%, pterossauro
%
%\item\label{rulegrp_tm} regra: \texttt{.t2m} -- tmese % [new rule] needs to be incorporated and checked
%
%\item\label{rulegrp_cza} regra: \texttt{c2za} -- czar
%
%\item\label{rulegrp_s2} regra: \texttt{.s2} -- stalinismo
%
%\item\label{rulegrp_t2} regra: \texttt{.t2} -- tsunami
%
%\item\label{rulegrp_mn} regra: \texttt{.m2n} -- mnemônico
%
%\item\label{rulegrp_zz} regra: \texttt{z1z}, \texttt{p1p}, \texttt{z2z.} -- pizza, shopping, jazz
%
%\item\label{rulegrp_sub} regra: \texttt{su2b3r}, \texttt{su2b3l} -- sublunar, subrotina \\
%exceção: \texttt{.su3b4li}
%%exceção: \texttt{su3b4li1nh}, \texttt{su3b4li1ma}, \texttt{su3b4li1me}, \texttt{su3b4li1mid} -- sublinhar, sublimar, sublime, sublimidade
%
%
%\item\label{rulegrp_so} regra: \texttt{1sô} -- subsônico
%
%\item\label{rulegrp_pseu} regra: \texttt{1p2seu1d} -- pseudônimo
%
%%\item\label{rulegrp_ein} regra: \texttt{e1in1c}, \texttt{e1in1f}, \texttt{e1in1g}, \texttt{e1in1s}, \texttt{e1in1t}, \texttt{e1in1v} -- reincidência, reinfecção, reingressa, reinserção, reintegração, reinventar \\
%%exceção: \texttt{e1in2st}, \texttt{e1in2sc} -- reinstalado, reinscrever
%
%\item\label{rulegrp_ein} regra: \texttt{e1inc}, \texttt{e1inf}, \texttt{e1ing}, \texttt{e1ins}, \texttt{e1int}, \texttt{e1inv} -- reincidência, reinfecção, reingressa, reinserção, reintegração, reinventar \\
%% exceção: \texttt{e1in2st}, \texttt{e1in2sc} -- reinstalado, reinscrever
%
%\item\label{rulegrp_unit} regra: \texttt{u1int}, \texttt{u1ind} -- contribuinte, excluindo
%
%\item\label{rulegrp_oin} regra: \texttt{o1in} -- coincidência, agroindustrial, bioinformática, amendoins, gastrointestinal
%exceção: \texttt{o1ina} -- boina
%
%\item\label{rulegrp_eim} regra: \texttt{e1imp} -- reimpresso%, coimbra, coimperador
%%exceção: \texttt{que2i}, \texttt{te2im} -- queimar, queimadura, teimosia, tirateima
%
%\item\label{rulegrp_aa} regra: \texttt{a1â}, \texttt{a1ã}, \texttt{a1é}, \texttt{a1í}, \texttt{a1ó}, \texttt{a1ô}, \texttt{a1ú}, \texttt{e1á}, \texttt{e1â}, \texttt{e1ã}, \texttt{e1é}, \texttt{e1ê}, \texttt{e1í}, \texttt{e1ó}, \texttt{e1ô}, \texttt{e1ú}, \texttt{é1o}, \texttt{i1á}, \texttt{i1ã}, \texttt{i1é}, \texttt{i1í}, \texttt{i1ó}, \texttt{i1u}, \texttt{i1ú}, \texttt{í1a}, \texttt{o1á}, \texttt{o1ã}, \texttt{o1é}, \texttt{o1ê}, \texttt{o1í}, \texttt{o1ó}, \texttt{u1á}, \texttt{u1ã}, \texttt{u1é}, \texttt{u1ê}, \texttt{u1í}, \texttt{ú1o} --  abraâmico, abraão, aéreo, país, caótico, faraônico, saúde, balneário, oceânico, campeã, feérico, veêm, veículo, teórico, napoleônico, conteúdo, néon, diário, região, soviético, iídiche, periódico, feiura, viúva, maníaco, razoável, joão, poético, boêmia, heroísmo, alcoólico, usuário, itapuã, suécia, cauê, suíça, flúor \\
%exceção: \texttt{1gu2é}, \texttt{1gu2ê}, \texttt{1gu2í}, \texttt{1qu2á}, \texttt{1qu2é}, \texttt{1qu2ê}, \texttt{1qu2í} -- alguém, português, linguística, aquático, inquérito, sequência, química
%% the inclusion of rules e1é, é1o, i1í and í1a cause problem and if exceptions are necessary
%
%\item\label{rulegrp_gua} regra: \texttt{1gu4á}, \texttt{1gu4ã}, \texttt{1qu4ã} -- jaraguá, saguão, quão
%
%\item\label{rulegrp_air} regra: \texttt{a1ir.}, \texttt{u1ir.} -- sair, diminuir
%
%\item\label{rulegrp_aiz} regra: \texttt{u1iz.}, \texttt{a1iz.} -- juiz, raiz
%
%\item\label{rulegrp_ia} regra: \texttt{í1a} -- baía
%
%\item\label{rulegrp_ain} regra: \texttt{a1ind}, \texttt{a1i1nh} -- ainda, rainha
%
%\item\label{rulegrp_oin} regra: \texttt{o1i1nh} -- moinho
%
%\item\label{rulegrp_aui} regra: \texttt{au1i1c}, \texttt{du1i1c}, \texttt{u1i1ç}, \texttt{u1i1d}, \texttt{cu2i}, \texttt{dru2i}, \texttt{flu2id}, \texttt{bu1i1n}, \texttt{cu1i1n}, \texttt{fu1i1n}, \texttt{nu1i1n}, \texttt{ru1i1na}, \texttt{ru1i1no}, \texttt{su1i1ti}, \texttt{tu1i1ti}, \texttt{u1iz} -- cacauicultor, sanduicheira, constituição, continuidade, cuidador, druida, fluido, tabuinha, picuinha, fuinha, genuinamente, arruinado, arruinou, jesuitismo, intuitivo, juizado \\ 
%exceção: \texttt{cu3i1da1de}, \texttt{bu1i2n1d}, \texttt{bu1i2n1t}  -- acuidade, contribuindo, contribuinte
%
%\item\label{rulegrp_iur} regra: \texttt{i1ur} -- diurno
%
%\item\label{rulegrp_io} regra: \texttt{í1o} -- íon
%
%\item\label{rulegrp_uo} regra: \texttt{ú1o} -- flúor
%
%\item\label{rulegrp_proi} regra: \texttt{pro1i1b} -- proibição
%
%\item\label{rulegrp_iun} regra: \texttt{i1un} -- triunfar
%
%\item\label{rulegrp_quo} regra: \texttt{1qu4ó}, \texttt{1qu4â} -- quórum, equânime
%
%\item\label{rulegrp_eo} regra: \texttt{é3o} -- alvéolo
%
%\item\label{rulegrp_ia} \st{regra: \texttt{i4a.}, \texttt{i4e.}, \texttt{i4o.}, \texttt{o4a.}, \texttt{u4a.} -- economia, espécie, vazio, destoa, institua}
%
%%\item\label{rulegrp_lo} regra: \texttt{1lô} -- camelô
%
%%\item\label{rulegrp_co} regra: \texttt{1cô} -- recôncavo
%
%\item\label{rulegrp_bo} regra: \texttt{1bô}, \texttt{1cô}, \texttt{1dô}, \texttt{1fô}, \texttt{1gô}, \texttt{1lô}, \texttt{1pô}, \texttt{1mô}, \texttt{1nô}, \texttt{1rô}, \texttt{1tô}, \texttt{1vô}, \texttt{1xô}, \texttt{1zô} -- robô, recôncavo, judô, telefônica, xangô, camelô, capô, sumô, econômico, tarô, chatô, vovô, saxônia, amazônia
%
%\item\label{rulegrp_cco} regra: \texttt{1çô} -- maçônico
%
%% As sequências gráficas ia, ie, io, oa, ua quando em posição
%% final átona, que são normalmente pronunciadas como ditongos
%% crescentes, mas que podem corresponder foneticamente a duas
%% vogais (hiato), sobretudo quando pronunciadas pausadamente:
%% vitó-ria, espé-cie, exercí-cio, nó-doa, lé-gua, tê-nue, ambí-guo.
%    \setcounter{numberRulesGroups}{\value{enumi}}
%\end{rules}
%
%The \NumberOfNewRules{} rules were grouped above in a list of
%\arabic{numberRulesGroups} types of rules. They may be further organized into four
%large groups. The first, which comprises rules \ref{rulegrp_psi} to
%\ref{rulegrp_zz}, includes consonant clusters such as \emph{czar}, \emph{ptose}
%and \emph{gnomo}. They, unlike the examples that will be exposed in section
%\ref{secxxx} (exceptions - item 9 currently), present a set of derived words,
%which makes their marking advantageous in view of the number of cases that are
%included in this marking rule. The second group, comprising rules
%\ref{rulegrp_sub} to \ref{rulegrp_eim}, delimits the morphological boundary
%between prefixes and radicals. As noted, although phonological issues guide the
%separation of numerous words in Portuguese, there are also those that are
%guided by morphology. This is the case of words that have the prefixes
%\emph{sub-} and \emph{re-}, such as \emph{sublunar} and \emph{reinserção}. The
%third group, comprising rules \ref{rulegrp_aa} to \ref{rulegrp_ia}, seeks to
%understand a set of words that have vowel combinations that do not follow the
%general rules. This is because the Portuguese language has vowel encounters
%with the second vowel graphically marked that can be separated, forming
%hiatuses, such as \emph{caótico}, \emph{balneário} and \emph{razoável}, while
%there are also words with a similar structure that constitute a diphthong, such
%as \emph{português}, \emph{alguém} and \emph{linguística}. It is remarkable, of
%course, that the latter are formed by the digraphs \emph{qu-} and \emph{gu-},
%while the former by vowels other than \emph{i} and \emph{u}. The fourth and
%last group, in turn, which comprises rules \ref{rulegrp_lo} to
%\ref{rulegrp_cco}, which are counterparts of rules that were already in the
%default rules, but did not contemplate the cases with certain accents. They
%were then added to encompass words such as \emph{camelô}, \emph{recôncavo},
%\emph{amazônia}, and \emph{maçônico}.
%
%It is important to highlight that the words included in these rules are, in general, of low frequency and were incorporated into the Portuguese language without a phonotactic adaptation, which causes these idiosyncrasies and exceptions to the language. They are characterized as marked cases, since it is not possible to defend that they reproduce a phonological pattern of the language because they are not productive, that is, they are not taken as an example and/or derive new words.
%
%
%O conjunto de \NumberOfNewRules{} regras é apresentado a seguir:
%\begin{multicols}{6}
%    \footnotesize
%    \verbatiminput{../data/patch.TeX.pt-br.patterns}
%\end{multicols}
%
%A Tabela \ref{tab-resultados} apresenta os erros cometidos no conjunto das
%37798 palavras analisadas. As palavras omitidas são aquelas que foram
%hifenizadas corretamente pelos dois conjuntos de regra. Utilizamos aqui a
%marcação adotada pelo \emph{Patgen}: \texttt{*} para indicar pontos de
%hifenização corretamente marcados, \texttt{.} para indicar pontos de hifenização
%marcados erroneamente e \texttt{-} para indicar pontos de hifenização não
%encontrados. Os pontos marcados erroneamente e aqueles não encontrados são
%considerados erros. Quando ocorre algum erro de hifenização, marcaremos a
%palavra com \xmark. Quando não houver erro algum, marcaremos a palavra com
%\cmark. Ao final, podemos verificar que as regras padrão apresentaram 1368 erros
%(3,62\%) e as novas regras 132 erros (0,35\%).
%
%\begin{longtable}{l l l}
%\caption{Comparativo dos erros cometidos pelos dois conjuntos de regras de
%hifenização.}\label{tab-resultados}\\
%    \hline
%    palavra & regras padrão & regras adicionais \\
%    \hline
%    \endhead
%    \input{comparedresults.tex}
%\end{longtable}
%

\printbibliography

\appendix
\section{Hyphenation results}
\input{appendix-table.tex}

\end{document}