### 2024-06-14 article/comparedresults-appprop.tex

abrupto & a.b-rup*to \xmark & a.b-rup*to \xmark \\
apneia & a-p.nei*a \xmark & a-p.nei-a \xmark \\
arábias & a*rá*bi.as \xmark & a*rá*bi.as \xmark \\
brechtiano & bre.ch*ti*a*no \xmark & bre.ch*ti*a*no \xmark \\
czarina & c.za*ri*na \xmark & cza*ri*na \cmark \\
czarismo & c.za*ris*mo \xmark & cza*ris*mo \cmark \\
czarista & c.za*ris*ta \xmark & cza*ris*ta \cmark \\
dispneia & dis-p.nei*a \xmark & dis-p.nei-a \xmark \\
dzeta & d.ze*ta \xmark & d.ze*ta \xmark \\
exéquias & e*xé*qui.as \xmark & e*xé*qui.as \xmark \\
gnaisse & g.nais*se \xmark & g.nais*se \xmark \\
gnomo & g.no*mo \xmark & gno*mo \cmark \\
gnose & g.no*se \xmark & gno*se \cmark \\
gnosiologia & g.no*si*o*lo*gi*a \xmark & gno*si*o*lo*gi-a \xmark \\
gnosticismo & g.nos*ti*cis*mo \xmark & gnos*ti*cis*mo \cmark \\
gnóstico & g.nós*ti*co \xmark & gnós*ti*co \cmark \\
hiperalgesia & hi*pe.r-al*ge*si*a \xmark & hi*pe.r-al*ge*si-a \xmark \\
mnemônica & m.ne-mô*ni*ca \xmark & mne*mô*ni*ca \cmark \\
mnemônico & m.ne-mô*ni*co \xmark & mne*mô*ni*co \cmark \\
neerlandês & ne.er*lan*dês \xmark & ne.er*lan*dês \xmark \\
núpcias & núp*ci.as \xmark & núp*ci.as \xmark \\
pneumática & p.neu*má*ti*ca \xmark & pneu*má*ti*ca \cmark \\
pneumático & p.neu*má*ti*co \xmark & pneu*má*ti*co \cmark \\
pneumatologia & p.neu*ma*to*lo*gi*a \xmark & pneu*ma*to*lo*gi-a \xmark \\
pneumococo & p.neu*mo*co*co \xmark & pneu*mo*co*co \cmark \\
pneumoconiose & p.neu*mo*co*ni*o*se \xmark & pneu*mo*co*ni*o*se \cmark \\
pneumogástrico & p.neu*mo*gás*tri*co \xmark & pneu*mo*gás*tri*co \cmark \\
pneumologia & p.neu*mo*lo*gi*a \xmark & pneu*mo*lo*gi-a \xmark \\
pneumonia & p.neu*mo*ni*a \xmark & pneu*mo*ni-a \xmark \\
pneumônico & p.neu-mô*ni*co \xmark & pneu*mô*ni*co \cmark \\
pneumotórax & p.neu*mo*tó*rax \xmark & pneu*mo*tó*rax \cmark \\
pneu & p.neu \xmark & pneu \cmark \\
pseudônimo & p.seu-dô*ni*mo \xmark & pseu*dô*ni*mo \cmark \\
psicanálise & p.si*ca*ná*li*se \xmark & psi*ca*ná*li*se \cmark \\
psicanalista & p.si*ca*na*lis*ta \xmark & psi*ca*na*lis*ta \cmark \\
psicanalítico & p.si*ca*na*lí*ti*co \xmark & psi*ca*na*lí*ti*co \cmark \\
psicodrama & p.si*co*dra*ma \xmark & psi*co*dra*ma \cmark \\
psicografia & p.si*co*gra*fi*a \xmark & psi*co*gra*fi-a \xmark \\
psicógrafo & p.si*có*gra*fo \xmark & psi*có*gra*fo \cmark \\
psicolinguista & p.si*co*lin*guis*ta \xmark & psi*co*lin*guis*ta \cmark \\
psicologia & p.si*co*lo*gi*a \xmark & psi*co*lo*gi-a \xmark \\
psicologicamente & p.si*co*lo*gi*ca*men*te \xmark & psi*co*lo*gi*ca*men*te \cmark \\
psicológico & p.si*co*ló*gi*co \xmark & psi*co*ló*gi*co \cmark \\
psicologismo & p.si*co*lo*gis*mo \xmark & psi*co*lo*gis*mo \cmark \\
psicologista & p.si*co*lo*gis*ta \xmark & psi*co*lo*gis*ta \cmark \\
psicologizar & p.si*co*lo*gi*zar \xmark & psi*co*lo*gi*zar \cmark \\
psicólogo & p.si*có*lo*go \xmark & psi*có*lo*go \cmark \\
psicométrico & p.si*co*mé*tri*co \xmark & psi*co*mé*tri*co \cmark \\
psicomotor & p.si*co*mo*tor \xmark & psi*co*mo*tor \cmark \\
psicomotricidade & p.si*co*mo*tri*ci*da*de \xmark & psi*co*mo*tri*ci*da*de \cmark \\
psicopata & p.si*co*pa*ta \xmark & psi*co*pa*ta \cmark \\
psicopático & p.si*co*pá*ti*co \xmark & psi*co*pá*ti*co \cmark \\
psicopatologia & p.si*co*pa*to*lo*gi*a \xmark & psi*co*pa*to*lo*gi-a \xmark \\
psicopatológico & p.si*co*pa*to*ló*gi*co \xmark & psi*co*pa*to*ló*gi*co \cmark \\
psicopedagogia & p.si*co*pe*da*go*gi*a \xmark & psi*co*pe*da*go*gi-a \xmark \\
psicopedagógico & p.si*co*pe*da*gó*gi*co \xmark & psi*co*pe*da*gó*gi*co \cmark \\
psicopedagogo & p.si*co*pe*da*go*go \xmark & psi*co*pe*da*go*go \cmark \\
psicose & p.si*co*se \xmark & psi*co*se \cmark \\
psicossexual & p.si*cos*se*xu*al \xmark & psi*cos*se*xu*al \cmark \\
psicossomático & p.si*cos*so*má*ti*co \xmark & psi*cos*so*má*ti*co \cmark \\
psicotécnico & p.si*co*téc*ni*co \xmark & psi*co*téc*ni*co \cmark \\
psicoterapeuta & p.si*co*te*ra*peu*ta \xmark & psi*co*te*ra*peu*ta \cmark \\
psicoterapia & p.si*co*te*ra*pi*a \xmark & psi*co*te*ra*pi-a \xmark \\
psicoterápico & p.si*co*te*rá*pi*co \xmark & psi*co*te*rá*pi*co \cmark \\
psicótico & p.si*có*ti*co \xmark & psi*có*ti*co \cmark \\
psicotrópico & p.si*co*tró*pi*co \xmark & psi*co*tró*pi*co \cmark \\
psique & p.si*que \xmark & psi*que \cmark \\
psiquiatra & p.si*qui*a*tra \xmark & psi*qui*a*tra \cmark \\
psiquiatria & p.si*qui*a*tri*a \xmark & psi*qui*a*tri-a \xmark \\
psiquiátrico & p.si*qui-á*tri*co \xmark & psi*qui*á*tri*co \cmark \\
psíquico & p.sí*qui*co \xmark & psí*qui*co \cmark \\
psiquismo & p.si*quis*mo \xmark & psi*quis*mo \cmark \\
psitaciforme & p.si*ta*ci*for*me \xmark & psi*ta*ci*for*me \cmark \\
psitacismo & p.si*ta*cis*mo \xmark & psi*ta*cis*mo \cmark \\
psoas & p.so*as \xmark & p.so*as \xmark \\
psoríase & p.so*rí-a*se \xmark & p.so*rí*a*se \xmark \\
pterossauro & p.te*ros*sau*ro \xmark & pte*ros*sau*ro \cmark \\
ptose & p.to*se \xmark & pto*se \cmark \\
quartzito & quar-t.zi*to \xmark & quar-t.zi*to \xmark \\
quartzo & quar-t.zo \xmark & quar-t.zo \xmark \\
stalinismo & s.ta*li*nis*mo \xmark & sta*li*nis*mo \cmark \\
stalinista & s.ta*li*nis*ta \xmark & sta*li*nis*ta \cmark \\
subliminar & su.b-li*mi*nar \xmark & su.b-li*mi*nar \xmark \\
sublingual & su.b-lin*gual \xmark & su.b-lin*gual \xmark \\
sublinha & su.b-li*nha \xmark & su.b-li*nha \xmark \\
subliteratura & su.b-li*te*ra*tu*ra \xmark & su.b-li*te*ra*tu*ra \xmark \\
sublocação & su.b-lo*ca*ção \xmark & sub*lo*ca*ção \cmark \\
sublocar & su.b-lo*car \xmark & sub*lo*car \cmark \\
sublunar & su.b-lu*nar \xmark & sub*lu*nar \cmark \\
tcheco & t.che*co \xmark & tche*co \cmark \\
tmese & t.me*se \xmark & tme*se \cmark \\
tzarista & t.za*ris*ta \xmark & tza*ris*ta \cmark \\