### 2024-06-14 article/comparedresults-sample.tex

ministério & mi*nis*té*ri.o \xmark & mi*nis*té*rio \cmark \\
próprio & pró*pri.o \xmark & pró*prio \cmark \\
salário & sa*lá*ri.o \xmark & sa*lá*rio \cmark \\
secretário & se*cre*tá*ri.o \xmark & se*cre*tá*rio \cmark \\
contrário & con*trá*ri.o \xmark & con*trá*rio \cmark \\
horário & ho*rá*ri.o \xmark & ho*rá*rio \cmark \\
necessário & ne*ces*sá*ri.o \xmark & ne*ces*sá*rio \cmark \\
relatório & re*la*tó*ri.o \xmark & re*la*tó*rio \cmark \\
adversário & ad*ver*sá*ri.o \xmark & ad*ver*sá*rio \cmark \\
cenário & ce*ná*ri.o \xmark & ce*ná*rio \cmark \\
diário & di-á*ri.o \xmark & di*á*rio \cmark \\
escritório & es*cri*tó*ri.o \xmark & es*cri*tó*rio \cmark \\
judiciário & ju*di*ci-á*ri.o \xmark & ju*di*ci*á*rio \cmark \\
equilíbrio & e*qui*lí*bri.o \xmark & e*qui*lí*brio \cmark \\
proprietário & pro*pri*e*tá*ri.o \xmark & pro*pri*e*tá*rio \cmark \\
aniversário & a*ni*ver*sá*ri.o \xmark & a*ni*ver*sá*rio \cmark \\
usuário & u*su-á*ri.o \xmark & u*su*á*rio \cmark \\
sério & sé*ri.o \xmark & sé*rio \cmark \\
território & ter*ri*tó*ri.o \xmark & ter*ri*tó*rio \cmark \\
seminário & se*mi*ná*ri.o \xmark & se*mi*ná*rio \cmark \\
auditório & au*di*tó*ri.o \xmark & au*di*tó*rio \cmark \\
laboratório & la*bo*ra*tó*ri.o \xmark & la*bo*ra*tó*rio \cmark \\
bancário & ban*cá*ri.o \xmark & ban*cá*rio \cmark \\
monetário & mo*ne*tá*ri.o \xmark & mo*ne*tá*rio \cmark \\
repertório & re*per*tó*ri.o \xmark & re*per*tó*rio \cmark \\
cemitério & ce*mi*té*ri.o \xmark & ce*mi*té*rio \cmark \\
compulsório & com*pul*só*ri.o \xmark & com*pul*só*rio \cmark \\
calendário & ca*len*dá*ri.o \xmark & ca*len*dá*rio \cmark \\
obrigatório & o*bri*ga*tó*ri.o \xmark & o*bri*ga*tó*rio \cmark \\
mistério & mis*té*ri.o \xmark & mis*té*rio \cmark \\
documentário & do*cu*men*tá*ri.o \xmark & do*cu*men*tá*rio \cmark \\
tributário & tri*bu*tá*ri.o \xmark & tri*bu*tá*rio \cmark \\
polícia & po*lí*ci.a \xmark & po*lí*cia \cmark \\
agência & a*gên*ci.a \xmark & a*gên*cia \cmark \\
presidência & pre*si*dên*ci.a \xmark & pre*si*dên*cia \cmark \\
violência & vi*o*lên*ci.a \xmark & vi*o*lên*cia \cmark \\
experiência & ex*pe*ri*ên*ci.a \xmark & ex*pe*ri*ên*cia \cmark \\
tendência & ten*dên*ci.a \xmark & ten*dên*cia \cmark \\
importância & im*por*tân*ci.a \xmark & im*por*tân*cia \cmark \\
denúncia & de*nún*ci.a \xmark & de*nún*cia \cmark \\
existência & e*xis*tên*ci.a \xmark & e*xis*tên*cia \cmark \\
ciência & ci*ên*ci.a \xmark & ci*ên*cia \cmark \\
ausência & au*sên*ci.a \xmark & au*sên*cia \cmark \\
emergência & e*mer*gên*ci.a \xmark & e*mer*gên*cia \cmark \\
referência & re*fe*rên*ci.a \xmark & re*fe*rên*cia \cmark \\
influência & in*flu-ên*ci.a \xmark & in*flu*ên*cia \cmark \\
conferência & con*fe*rên*ci.a \xmark & con*fe*rên*cia \cmark \\
notícia & no*tí*ci.a \xmark & no*tí*cia \cmark \\
consequência & con*se-quên*ci.a \xmark & con*se*quên*cia \cmark \\
assistência & as*sis*tên*ci.a \xmark & as*sis*tên*cia \cmark \\
distância & dis*tân*ci.a \xmark & dis*tân*cia \cmark \\
resistência & re*sis*tên*ci.a \xmark & re*sis*tên*cia \cmark \\
consciência & cons*ci*ên*ci.a \xmark & cons*ci*ên*cia \cmark \\
eficiência & e*fi*ci*ên*ci.a \xmark & e*fi*ci*ên*cia \cmark \\
infância & in*fân*ci.a \xmark & in*fân*cia \cmark \\
inteligência & in*te*li*gên*ci.a \xmark & in*te*li*gên*cia \cmark \\
vigilância & vi*gi*lân*ci.a \xmark & vi*gi*lân*cia \cmark \\
competência & com*pe*tên*ci.a \xmark & com*pe*tên*cia \cmark \\
independência & in*de*pen*dên*ci.a \xmark & in*de*pen*dên*cia \cmark \\
início & i*ní*ci.o \xmark & i*ní*cio \cmark \\
comércio & co*mér*ci.o \xmark & co*mér*cio \cmark \\
negócio & ne*gó*ci.o \xmark & ne*gó*cio \cmark \\
palácio & pa*lá*ci.o \xmark & pa*lá*cio \cmark \\
anúncio & a*nún*ci.o \xmark & a*nún*cio \cmark \\
márcio & már*ci.o \xmark & már*cio \cmark \\
benefício & be*ne*fí*ci.o \xmark & be*ne*fí*cio \cmark \\
história & his*tó*ri.a \xmark & his*tó*ria \cmark \\
vitória & vi*tó*ri.a \xmark & vi*tó*ria \cmark \\
indústria & in*dús*tri.a \xmark & in*dús*tria \cmark \\
memória & me*mó*ri.a \xmark & me*mó*ria \cmark \\
miséria & mi*sé*ri.a \xmark & mi*sé*ria \cmark \\
secretária & se*cre*tá*ri.a \xmark & se*cre*tá*ria \cmark \\
médio & mé*di.o \xmark & mé*dio \cmark \\
rádio & rá*di.o \xmark & rá*dio \cmark \\
incêndio & in*cên*di.o \xmark & in*cên*dio \cmark \\
estúdio & es*tú*di.o \xmark & es*tú*dio \cmark \\
índio & ín*di.o \xmark & ín*dio \cmark \\
psicologia & p.si*co*lo*gi*a \xmark & psi*co*lo*gi-a \xmark \\
psicológico & p.si*co*ló*gi*co \xmark & psi*co*ló*gi*co \cmark \\
psiquiatra & p.si*qui*a*tra \xmark & psi*qui*a*tra \cmark \\
psicanalista & p.si*ca*na*lis*ta \xmark & psi*ca*na*lis*ta \cmark \\
psicólogo & p.si*có*lo*go \xmark & psi*có*lo*go \cmark \\
domínio & do*mí*ni.o \xmark & do*mí*nio \cmark \\
raciocínio & ra*ci*o*cí*ni.o \xmark & ra*ci*o*cí*nio \cmark \\
condomínio & con*do*mí*ni.o \xmark & con*do*mí*nio \cmark \\
alumínio & a*lu*mí*ni.o \xmark & a*lu*mí*nio \cmark \\
hélio & hé*li.o \xmark & hé*lio \cmark \\
monopólio & mo*no*pó*li.o \xmark & mo*no*pó*lio \cmark \\
auxílio & au*xí*li.o \xmark & au*xí*lio \cmark \\
colégio & co*lé*gi.o \xmark & co*lé*gio \cmark \\
prestígio & pres*tí*gi.o \xmark & pres*tí*gio \cmark \\
relógio & re*ló*gi.o \xmark & re*ló*gio \cmark \\
câmbio & câm*bi.o \xmark & câm*bio \cmark \\
intercâmbio & in*ter*câm*bi.o \xmark & in*ter*câm*bio \cmark \\
subúrbio & su*búr*bi.o \xmark & su*búr*bio \cmark \\
média & mé*di.a \xmark & mé*dia \cmark \\
mídia & mí*di.a \xmark & mí*dia \cmark \\
município & mu*ni*cí*pi.o \xmark & mu*ni*cí*pio \cmark \\
princípio & prin*cí*pi.o \xmark & prin*cí*pio \cmark \\
contemporâneo & con*tem*po*râ*ne.o \xmark & con*tem*po*râ*neo \cmark \\
mediterrâneo & me*di*ter*râ*ne.o \xmark & me*di*ter*râ*neo \cmark \\
autópsia & au*tóp*si.a \xmark & au*tóp*sia \cmark \\
eutanásia & eu*ta*ná*si.a \xmark & eu*ta*ná*sia \cmark \\
vídeo & ví*de.o \xmark & ví*deo \cmark \\
hominídeo & ho*mi*ní*de.o \xmark & ho*mi*ní*deo \cmark \\
alívio & a*lí*vi.o \xmark & a*lí*vio \cmark \\
sérvio & sér*vi.o \xmark & sér*vio \cmark \\
aéreo & a-é*re.o \xmark & a*é*reo \cmark \\
térreo & tér*re.o \xmark & tér*reo \cmark \\
pneumonia & p.neu*mo*ni*a \xmark & pneu*mo*ni-a \xmark \\
pneumática & p.neu*má*ti*ca \xmark & pneu*má*ti*ca \cmark \\
represália & re*pre*sá*li.a \xmark & re*pre*sá*lia \cmark \\
sandália & san*dá*li.a \xmark & san*dá*lia \cmark \\
ginásio & gi*ná*si.o \xmark & gi*ná*sio \cmark \\
múmia & mú*mi.a \xmark & mú*mia \cmark \\
hérnia & hér*ni.a \xmark & hér*nia \cmark \\
exímio & e*xí*mi.o \xmark & e*xí*mio \cmark \\
crustáceo & crus*tá*ce.o \xmark & crus*tá*ceo \cmark \\
sublunar & su.b-lu*nar \xmark & sub*lu*nar \cmark \\
sítio & sí*ti.o \xmark & sí*tio \cmark \\
equilíbrio & e*qui*lí*bri.o \xmark & e*qui*lí*brio \cmark \\
rodízio & ro*dí*zi.o \xmark & ro*dí*zio \cmark \\
arbítrio & ar*bí*tri.o \xmark & ar*bí*trio \cmark \\
angústia & an*gús*ti.a \xmark & an*gús*tia \cmark \\
colóquio & co*ló*qui.o \xmark & co*ló*quio \cmark \\
cópia & có*pi.a \xmark & có*pia \cmark \\
espécie & es*pé*ci.e \xmark & es*pé*cie \cmark \\
série & sé*ri.e \xmark & sé*rie \cmark \\
sanguínea & san*guí*ne.a \xmark & san*guí*nea \cmark \\
petróleo & pe*tró*le.o \xmark & pe*tró*leo \cmark \\
vácuo & vá*cu.o \xmark & vá*cuo \cmark \\
lascívia & las*cí*vi.a \xmark & las*cí*via \cmark \\
indústria & in*dús*tri.a \xmark & in*dús*tria \cmark \\
área & á*re.a \xmark & á*rea \cmark \\
exéquias & e*xé*qui.as \xmark & e*xé*qui.as \xmark \\
gnomo & g.no*mo \xmark & gno*mo \cmark \\
estratégia & es*tra*té*gi.a \xmark & es*tra*té*gia \cmark \\
epitáfio & e*pi*tá*fi.o \xmark & e*pi*tá*fio \cmark \\
máfia & má*fi.a \xmark & má*fia \cmark \\
assíduo & as*sí*du.o \xmark & as*sí*duo \cmark \\
czarina & c.za*ri*na \xmark & cza*ri*na \cmark \\
cucurbitácea & cu*cur*bi*tá*ce.a \xmark & cu*cur*bi*tá*cea \cmark \\
tíbia & tí*bi.a \xmark & tí*bia \cmark \\
pétreo & pé*tre.o \xmark & pé*treo \cmark \\
teleósteo & te*le-ós*te.o \xmark & te*le*ós*teo \cmark \\
stalinismo & s.ta*li*nis*mo \xmark & sta*li*nis*mo \cmark \\
ósseo & ós*se.o \xmark & ós*seo \cmark \\
psoríase & p.so*rí-a*se \xmark & p.so*rí*a*se \xmark \\
próprio & pró*pri.o \xmark & pró*prio \cmark \\
apneia & a-p.nei*a \xmark & a-p.nei-a \xmark \\
mnemônica & m.ne-mô*ni*ca \xmark & mne*mô*ni*ca \cmark \\
imbróglio & im*bró*gli.o \xmark & im*bró*glio \cmark \\
laríngeo & la*rín*ge.o \xmark & la*rín*geo \cmark \\
supérfluo & su*pér*flu.o \xmark & su*pér*fluo \cmark \\
núcleo & nú*cle.o \xmark & nú*cleo \cmark \\
dúzia & dú*zi.a \xmark & dú*zia \cmark \\
várzea & vár*ze.a \xmark & vár*zea \cmark \\
névoa & né*vo.a \xmark & né*voa \cmark \\
gávea & gá*ve.a \xmark & gá*vea \cmark \\
quartzo & quar-t.zo \xmark & quar-t.zo \xmark \\
quartzito & quar-t.zi*to \xmark & quar-t.zi*to \xmark \\
tzarista & t.za*ris*ta \xmark & tza*ris*ta \cmark \\
mútuo & mú*tu.o \xmark & mú*tuo \cmark \\
estátua & es*tá*tu.a \xmark & es*tá*tua \cmark \\
mênstruo & mêns*tru.o \xmark & mêns*truo \cmark \\
tmese & t.me*se \xmark & tme*se \cmark \\
abrótea & a*bró*te.a \xmark & a*bró*tea \cmark \\
tcheco & t.che*co \xmark & tche*co \cmark \\
exéquias & e*xé*qui.as \xmark & e*xé*qui.as \xmark \\
precípuo & pre*cí*pu.o \xmark & pre*cí*puo \cmark \\
ptose & p.to*se \xmark & pto*se \cmark \\
pterossauro & p.te*ros*sau*ro \xmark & pte*ros*sau*ro \cmark \\
psíquico & p.sí*qui*co \xmark & psí*qui*co \cmark \\
pseudônimo & p.seu-dô*ni*mo \xmark & pseu*dô*ni*mo \cmark \\
hiperalgesia & hi*pe.r-al*ge*si*a \xmark & hi*pe.r-al*ge*si-a \xmark \\
contínuo & con*tí*nu.o \xmark & con*tí*nuo \cmark \\
contínua & con*tí*nu.a \xmark & con*tí*nua \cmark \\
neerlandês & ne.er*lan*dês \xmark & ne.er*lan*dês \xmark \\
fêmea & fê*me.a \xmark & fê*mea \cmark \\
águia & á*gui.a \xmark & á*guia \cmark \\
mágoa & má*go.a \xmark & má*goa \cmark \\
gnóstico & g.nós*ti*co \xmark & gnós*ti*co \cmark \\
gnosticismo & g.nos*ti*cis*mo \xmark & gnos*ti*cis*mo \cmark \\
gnaisse & g.nais*se \xmark & g.nais*se \xmark \\
efígie & e*fí*gi.e \xmark & e*fí*gie \cmark \\
drágea & drá*ge.a \xmark & drá*gea \cmark \\
dzeta & d.ze*ta \xmark & d.ze*ta \xmark \\
mitocôndria & mi*to-côn*dri.a \xmark & mi*to*côn*dria \cmark \\
amêndoa & a*mên*do.a \xmark & a*mên*doa \cmark \\
côdea & cô*de.a \xmark & cô*dea \cmark \\
cnidário & c.ni*dá*ri.o \xmark & c.ni*dá*rio \xmark \\
núpcias & núp*ci.as \xmark & núp*ci.as \xmark \\
quíchua & quí*chu.a \xmark & quí*chua \cmark \\
fímbria & fím*bri.a \xmark & fím*bria \cmark \\
brechtiano & bre.ch*ti*a*no \xmark & bre.ch*ti*a*no \xmark \\
arábias & a*rá*bi.as \xmark & a*rá*bi.as \xmark \\
plúmbeo & plúm*be.o \xmark & plúm*beo \cmark \\
abrupto & a.b-rup*to \xmark & a.b-rup*to \xmark \\