### 2024-06-14 article/resultsdefaultrules.tex

\section{Results for the default rules}

In this section we will apprise the performance of default \TeX{} hyphenation
rules on the list of hyphenated words, which creation was described in
\Cref{sec-dictionay}. \Cref{tab-default-results} summaries the results, showing
the number of words correctly hyphenated words by the default rules, the number
of words incorrectly hyphenated and the number of words where a hyphenation point
was missed.

\begin{table}[htbp]
\centering
\caption{Results of \TeX{} default hyphenation rules on \texttt{hyphenations6}, \texttt{hyphenations5} and \texttt{hyphenations4} dictionaries.}
\label{tab-default-results}
\begin{tabular}{lrrrr}
\toprule
word list 	& \# correct 		& \# incorrect		& \# missing 		& \# entries 		\\
\midrule
hyphenations6 	& \NumberOfCorrectSix{}	& \NumberOfWrongSix{}	& \NumberOfMissingSix{}	& \NumberOfTotalSix{} 	\\
hyphenations5 	& \NumberOfCorrectFive{}& \NumberOfWrongFive{}	& \NumberOfMissingFive{}& \NumberOfTotalFive{} 	\\
hyphenations4 	& \NumberOfCorrectFour{}& \NumberOfWrongFour{}	& \NumberOfMissingFour{}& \NumberOfTotalFour{} 	\\
total		& \NumberOfCorrectSum{} & \NumberOfWrongSum{} 	& \NumberOfMissingSum{}	& \NumberOfTotalSum{}   \\
\bottomrule
\end{tabular}
\end{table}

Each word in the list is hyphenated using the set of default rules using the \texttt{gohyphen} script
(the \texttt{gohyphenfull} script is used to see in the interplay among all rules that apply to a given word) \parencite{gundlach2021}.
The hyphenation resulted from the given rules is compared to the hyphenation in our dictionary, so
we may check if there were erros (incorrect or missing hyphenation points).